---
title: "01_dataset_creation"
author: "Álvaro Ritoré"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
# output:
#   html_notebook:
#     code_folding: hide
#     number_sections: yes
#     theme: flatly
#     toc: yes
#     toc_float: yes
---

# Environment

```{r message=FALSE, warning=FALSE}
library(bigrquery)
library(DBI)
library(dplyr)
library(summarytools)
library(sqldf)
library(tidyverse)
library(nlme)
library(zoo)
library(tableone)
library(readxl)
library(feather)
library(magrittr)
library(MatchIt)
library(kableExtra)
library(ggplot2)
library(naniar)
library(modeest)
```

# BigQuery related functions

This chunks creates the run_query and getSQL function.

```{r}

# Function that takes in a SQL command and runs it on BigQuery. This avoids the connection data in each iteration
run_query<-function(query){
  query_output<-dbGetQuery(con,query)
  return(query_output)
}

# Function for reading SQL files from a folder
getSQL <- function(filepath) {
  con = file(filepath, "r")
  sql.string <- ""
  while (TRUE) {
    line <- readLines(con, n = 1)
    if (length(line) == 0) {
      break
    }
    line <- gsub("\\t", " ", line)
    if (grepl("--", line) == TRUE) {
      line <- paste(sub("--", "/*", line), "*/")
    }
    sql.string <- paste(sql.string, line)
  }
  close(con)
  return(sql.string)
}
```

# Data Load

Loading queries and extracting the data

## Prolonged Mechanical Ventilation

```{r}
# Criterion of prolonged mechanical ventilation: highest number of consecutive days of mechanical ventilation six or more hours per day
vm_consecutive_days <- run_query(getSQL("sql/eICU/vm_consecutive_days.sql"))

# Criterion of prolonged mechanical ventilation: total consecutive or non-consecutive days of mechanical ventilation for six or more hours per day
vm_total_days <- run_query(getSQL("sql/vm_total_days.sql"))

# Criterion of prolonged mechanical ventilation: 96 hour ventilation + tracheotomy
vm_and_traq <- run_query(getSQL("sql/vm_and_traq.sql"))
```

## Traumatisms

```{r}
#Traumatic Brain Injury
tbi <- run_query(getSQL("sql/tbi.sql"))
tbi %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    tbi_seq1=max(tbi_seq_1), 
    tbi_anyseq=max(tbi_any_seq))

# Data cleaning to group by hadm_id
#General traumatisms
trauma_general <- run_query(getSQL("sql/trauma_general.sql"))
trauma_general %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    traumatic_general_seq_1=max(traumatic_general_seq_1), 
    traumatic_general_any_seq=max(traumatic_general_any_seq))

#Fractures
trauma_fractures <- run_query(getSQL("sql/trauma_fractures.sql"))
trauma_fractures %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    fractures_seq_1=max(fractures_seq_1),
    fractures_any_seq=max(fractures_any_seq))

#Luxations
trauma_luxations <- run_query(getSQL("sql/trauma_luxations.sql"))
trauma_luxations %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    luxation_seq_1=max(luxation_seq_1),
    luxation_any_seq=max(luxation_any_seq))

#Sprains
trauma_sprains <- run_query(getSQL("sql/trauma_sprains.sql"))
trauma_sprains %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    sprains_seq_1=max(sprains_seq_1),
    sprains_any_seq=max(sprains_any_seq))

#Burns
trauma_burns <- run_query(getSQL("sql/trauma_burns.sql"))
trauma_burns %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    burns_seq_1=max(burns_seq_1),
    burns_any_seq=max(burns_any_seq))

#Contusion
trauma_contusion <- run_query(getSQL("sql/trauma_contusions.sql"))
trauma_contusion %<>%
  group_by(subject_id, hadm_id) %>%
  summarize(
    contusion_seq_1=max(contusion_seq_1),
    contusion_any_seq=max(contusion_any_seq))

```

## Demographic information

```{r}
#Patients
patients <- run_query(getSQL("sql/patients.sql"))

#Admissions
admissions <- run_query(getSQL("sql/admissions.sql"))

#Stays in ICU
icustays <- run_query(getSQL("sql/icustays.sql"))
```

## Severity scores

```{r}
#Acute Physiology Score (APS)
apsiii <- run_query(getSQL("sql/apsiii.sql"))

#Simplified Acute Physiology Score (SAPS)
sapsii <- run_query(getSQL("sql/sapsii.sql"))

#Glasgow Coma Scale (GCS)
first_day_gcs <- run_query(getSQL("sql/first_day_gcs.sql"))

#Sequential Organ Failure Assessment (SOFA)
first_day_sofa <- run_query(getSQL("sql/first_day_sofa.sql"))
```

# Creation of Final Dataset

## Joining across datasets

```{r}
# Join patients tables by subject_id
dataset_join <- icustays %>%
  left_join(patients, by = "subject_id")

# Join tables by subject_id and hadm_id  
dataset_join <- Reduce(function(...) merge(..., by = c("subject_id", "hadm_id"), all.x=TRUE, sort = TRUE), list(
  dataset_join, 
  admissions, 
  tbi,
  trauma_general, 
  trauma_fractures, 
  trauma_luxations, 
  trauma_sprains, 
  trauma_burns, 
  trauma_contusion))
  
# Join tables by subject_id, hadm_id and stay_id (except first_day_gcs as it doesn't include hadm_id)
dataset_join <- Reduce(function(...) merge(..., by = c("subject_id","hadm_id", "stay_id"), all.x=TRUE, sort = TRUE), list(
  dataset_join,
  vm_consecutive_days, 
  vm_total_days,
  vm_and_traq,
  apsiii,
  sapsii,
  first_day_sofa
))

#Join first_day_gcs table
dataset_join %<>%
  left_join(first_day_gcs, by = c("subject_id", "stay_id"))

```

## Creation of new variables and adjustment

```{r}
# Convert columns names to lowercase
names(dataset_join) <- tolower(names(dataset_join))

# Rename columns for self-explanation purposes
total_dataset <- dataset_join %>%
  rename(
    "age" = anchor_age,
    "hosp_admittime" = admittime, #date and time the patient was admitted to the hospital
    "icu_admittime" = intime, #date and time the patient was transferred into the ICU
    "icu_outtime" = outtime, #date and time the patient was transferred out of the ICU
    "inhospital_death" = hospital_expire_flag, #patient death within the given hospitalization
    "date_of_death" = dod, #date of death for the patient (out of hospital mortality is available for MIMIC-IV patients up to one year post-hospital discharge)
  )

# Convert character-type variables to factor 
total_dataset %<>%
  mutate("gender" = as.factor(gender))

# Convert time/date variables in proper format 
total_dataset %<>%
  mutate(
    icu_outtime = as.POSIXct(as.character(icu_outtime), format = "%Y-%m-%d %H:%M:%S"),
    date_of_death = as.Date(as.character(date_of_death), format = "%Y-%m-%d"),
    inicu_death = as.integer(ifelse(difftime(as.Date(icu_outtime), date_of_death, units = "days") >= 0, 1, 0)))

# Variables:
## hospital_stay_days: length of stay for the given hospital stay
## icu_stay_days: length of stay for the patient for the given ICU stay
total_dataset %<>%
  mutate(hospital_stay_days = round(as.numeric((difftime(dischtime, hosp_admittime, units = "hours"))/24), 2),
         icu_stay_days = round(los, 2))

# Variables death_30days (mortality at 30 days) and death_year (mortality at 1 year)
total_dataset %<>%
    mutate(
    death_30days = as.integer(ifelse(difftime(date_of_death, hosp_admittime, units = "days") <= 30, 1, 0)),
    death_year = as.integer(ifelse(difftime(date_of_death, hosp_admittime, units = "days") <= 365, 1, 0)
    ))

# Traumatism variables
total_dataset %<>%
  mutate(
    anytrauma_seq1 = as.integer(ifelse(tbi_seq1 == 1 | contusion_seq_1 == 1 | fractures_seq_1 == 1 | traumatic_general_seq_1 == 1 | luxation_seq_1 == 1 | sprain_seq_1 == 1, 1, 0)),          
    anytrauma_anyseq = as.integer(ifelse(tbi_anyseq == 1 | contusion_any_seq == 1 | fractures_any_seq == 1 | traumatic_general_any_seq == 1 | luxation_any_seq == 1 | sprain_any_seq == 1, 1, 0)),
            )

# Variables mv_14days (prolonged mechanical ventilation, criterion of 14 days) and mv_96traq (prolonged mechanical ventilation, criterion of 96h of ventilation + traqueostomy)
total_dataset %<>%
  mutate(
    mv_consecutive_7days = as.integer(ifelse(consecutive_vm_days>=7, 1, 0)),
    mv_consecutive_14days = as.integer(ifelse(consecutive_vm_days>=14, 1, 0)),
    mv_consecutive_21days = as.integer(ifelse(consecutive_vm_days>=21, 1, 0)),
    mv_total_7days = as.integer(ifelse(total_vm_days>=7, 1, 0)),
    mv_total_14days = as.integer(ifelse(total_vm_days>=14, 1, 0)),
    mv_total_21days = as.integer(ifelse(total_vm_days>=21, 1, 0)),
    mv_96traq = as.integer(ifelse(!is.na(trach_time), 1, 0)),
    prolonged_mec_vent = as.integer(ifelse((mv_consecutive_14days == 1 |
                                             mv_96traq == 1), 1, 0
    ))
  )

# Gravity scores variables
total_dataset %<>%
  mutate(
    apsiii_prob = round(apsiii_prob*100, 2), 
    sapsii_prob = round(sapsii_prob*100, 2)
  )

```

## Selection of required columns

```{r}

# Select the desired columns and order the result
total_dataset %<>% 
  select(subject_id, hadm_id, stay_id, age, gender, icu_admittime, hospital_stay_days, icu_stay_days, discharge_location, inhospital_death, inicu_death, death_30days, death_year, tbi_seq1, tbi_anyseq, anytrauma_seq1, anytrauma_anyseq, consecutive_vm_days, mv_consecutive_7days, mv_consecutive_14days, mv_consecutive_21days, total_vm_days, mv_total_7days, mv_total_14days, mv_total_21days, mv_96traq, prolonged_mec_vent, apsiii, apsiii_prob, sapsii, sapsii_prob, gcs_min, gcs_motor, gcs_verbal, gcs_eyes, gcs_unable, sofa, respiration, coagulation, liver, cardiovascular, cns, renal) %>%
  arrange(subject_id, icu_admittime)

```

## Addressing missingness

```{r}
# Map missingness in total_dataset before NA filling  
##Columns with na values
print('Columns in total_dataset having NAs')
names(total_dataset)[colSums(is.na(total_dataset)) > 0]

print('Nº of columns with na values')
n_var_miss(total_dataset)

##Plot columns and show descriptive table with NA values
miss_summary_preclean <- miss_var_summary(total_dataset)
miss_summary_preclean
gg_miss_var(total_dataset, show_pct=TRUE)
  
```
## NA filling
```{r}
# Columns NA filling
total_dataset_clean <- total_dataset %>%
  mutate(
##Column discharge_location is a non-dichotomous categorical variable. When having an NA and the patient died in-hospital, a "DIED" is filled. The other NAs are filled with the most frequent value (mode) using the mode.
    discharge_location = ifelse(!is.na(discharge_location) & inhospital_death == 1, "DIED", discharge_location), 
    discharge_location = ifelse(is.na(discharge_location), modeest::mfv(discharge_location), discharge_location),
## Discontinuous variables related with severity scores: 
### Glascow Coma Scale (GCS) 
### All rows in MIMIC-IV with gcs_unable=1 also have gcs_verbal=0 (15287 rows). There are also 95 rows with gcs_unable=0 with gcs_verbal=0 (95 rows). Since all sub gcs variables must display a minimal value equal to 1, this is corrected. 
    gcs_verbal = as.integer(ifelse(gcs_verbal==0, 1, gcs_verbal)),
### gcs_motor, gcs_verbal, gcs_eyes missing values are filled with the mode
    gcs_verbal = as.integer(ifelse(is.na(gcs_verbal), modeest::mfv(gcs_verbal, na_rm=TRUE), gcs_verbal)),
    gcs_motor = as.integer(ifelse(is.na(gcs_motor), modeest::mfv(gcs_motor, na_rm=TRUE), gcs_motor)),
    gcs_eyes = as.integer(ifelse(is.na(gcs_eyes), modeest::mfv(gcs_eyes, na_rm=TRUE), gcs_eyes)),
# Fill the NA in tbi_seq1 and tbi_anyseq to avoid errors in next lines
    tbi_seq1 = ifelse(is.na(tbi_seq1), 0, tbi_seq1),
    tbi_anyseq = ifelse(is.na(tbi_anyseq), 0, tbi_anyseq),
### All 3 gcs variables are added together to obtain gcs_min, only when gcs_min is NA (530 rows, also with gcs_unable missing), whose gcs_min is filled completely based on mode; and patients with TBI and gcs_unable=1. Rows with gcs_unable=1 and not TBI has to maintain their 15 value. 
    gcs_min = as.integer(ifelse(is.na(gcs_min) | (tbi_anyseq==1 & gcs_unable==1), gcs_motor + gcs_verbal + gcs_eyes, gcs_min)),
### SOFA: respiration, coagulation, liver, cardiovascular, cns and renal variables are filled with the median, and those 6 are added together to obtain SOFA
    respiration = ifelse(is.na(respiration), median(respiration, na_rm=TRUE), respiration), 
    coagulation = ifelse(is.na(coagulation), median(coagulation, na_rm=TRUE), coagulation), 
    liver = ifelse(is.na(liver), median(liver, na_rm=TRUE), liver), 
    cardiovascular = ifelse(is.na(cardiovascular), median(cardiovascular, na_rm=TRUE), cardiovascular), 
    cns = ifelse(is.na(cns), median(cns, na_rm=TRUE), cns), 
    renal = ifelse(is.na(renal), median(renal, na_rm=TRUE), renal), 
    sofa = as.integer(respiration + coagulation + liver + cardiovascular + cns + renal)) %>%
##The rest of the NA-containing columns are filled with a '0' value. These include dichotomous variables (inicu_death, death_30days, death_year, tbi_seq1, tbi_anyseq, consecutive_vm_days, mv_consecutive_7days, mv_consecutive_14days, mv_consecutive_21days, mv_total_7days, mv_total_14days, mv_total_21days and prolonged_mec_vent), as well as discontinuous variables related with ventilation (consecutive_vm_days and total_vm_days)
  mutate_at(vars(10:ncol(total_dataset)), ~ifelse(is.na(.), 0, .))

```

## Confirm 0% missingness
```{r}
# Map missingness in total_dataset_clean after NA filling  
##Columns with na values
print('Columns in total_dataset_clean having NAs')
names(total_dataset_clean)[colSums(is.na(total_dataset_clean)) > 0]

print('Nº of columns with na values')
n_var_miss(total_dataset_clean)

##Plot columns and show descriptive table with NA values
miss_summary_postclean <- miss_var_summary(total_dataset_clean)
miss_summary_postclean
gg_miss_var(total_dataset_clean, show_pct=TRUE)
  
```
## Creation of additional variables
```{r}
# Create variable tbi severe and non-severe tbi, as well as trauma_type, containing 4 types of trauma in our dataset: TBI, general traumatism without TBI, severe TBI, non-severe TBI, for both seq1 and anyseq 
total_dataset_clean %<>%
  mutate(
    tbi_severe_seq1 = ifelse(tbi_seq1 == 1 & gcs_min <= 8, 1, 0),
    tbi_nonsevere_seq1 = ifelse(tbi_seq1 == 1 &
                                  gcs_min > 8, 1, 0),
    trauma_type_seq1 = as.factor(
      ifelse(anytrauma_seq1 == 0, "non_trauma",
             ifelse(anytrauma_seq1 == 1 & tbi_seq1 == 0, "trauma_non_tbi",
                    ifelse(tbi_severe_seq1 == 1, "tbi_severe",
                           ifelse(tbi_nonsevere_seq1 == 1, "tbi_non_severe", NA)))
      )
    ),
    tbi_severe_anyseq = ifelse(tbi_anyseq == 1 & gcs_min <= 8, 1, 0),
    tbi_nonsevere_anyseq = ifelse(tbi_anyseq == 1 &
                                  gcs_min > 8, 1, 0),
    trauma_type_anyseq = as.factor(
      ifelse(anytrauma_anyseq == 0, "non_trauma",
             ifelse(anytrauma_anyseq == 1 & tbi_anyseq == 0, "trauma_non_tbi",
                    ifelse(tbi_severe_anyseq == 1, "tbi_severe",
                           ifelse(tbi_nonsevere_anyseq == 1, "tbi_non_severe", NA)))
      )
    )
  )

# Select and order columns
total_dataset_clean %<>% 
  select(subject_id:anytrauma_anyseq, tbi_severe_seq1, tbi_nonsevere_seq1, trauma_type_seq1, tbi_severe_anyseq, tbi_nonsevere_anyseq, trauma_type_anyseq, consecutive_vm_days:renal) %>%
  arrange(subject_id, icu_admittime)
```


## Final Dataset
```{r}
# Group all stay_id within joint hospital admissions

#Summarizing variables into hadm_id groups
group_dataset_1 <- total_dataset_clean %>%
  group_by(subject_id, hadm_id) %>%
  arrange (icu_admittime) %>% 
  summarize(
    age=min(age), 
    gender=first(gender), 
    icu_admittime=min(icu_admittime), 
    hospital_stay_days=first(hospital_stay_days),
    icu_stay_days=sum(icu_stay_days), 
    discharge_location=last(discharge_location),
    inhospital_death=max(inhospital_death), 
    inicu_death=max(inicu_death), 
    death_30days=max(death_30days), 
    death_year=max(death_year), 
    tbi_seq1=first(tbi_seq1), 
    tbi_anyseq=first(tbi_anyseq), 
    anytrauma_seq1=first(anytrauma_seq1),
    anytrauma_anyseq=first(anytrauma_anyseq),
    tbi_severe_seq1=first(tbi_severe_seq1), 
    tbi_nonsevere_seq1=first(tbi_nonsevere_seq1),
    trauma_type_seq1=first(trauma_type_seq1),
    tbi_severe_anyseq=first(tbi_severe_anyseq), 
    tbi_nonsevere_anyseq=first(tbi_nonsevere_anyseq),
    trauma_type_anyseq=first(trauma_type_anyseq),
    consecutive_vm_days=max(consecutive_vm_days), 
    mv_consecutive_7days=max(mv_consecutive_7days), 
    mv_consecutive_14days=max(mv_consecutive_14days),
    mv_consecutive_21days=max(mv_consecutive_21days),
    total_vm_days=sum(total_vm_days), 
    mv_total_7days = ifelse(total_vm_days>=7, 1, 0),
    mv_total_14days = ifelse(total_vm_days>=14, 1, 0),
    mv_total_21days = ifelse(total_vm_days>=21, 1, 0),
    mv_96traq=max(mv_96traq), 
    prolonged_mec_vent=max(prolonged_mec_vent), 
          ) %>%
    arrange(subject_id)

#Select severity scores related to first stay_id. It must be noted that stay_id in MIMIC-IV are not necessarily ordered numerically, so we order by ICU admission time.  
group_dataset_2 <- total_dataset_clean %>%
  group_by(subject_id, hadm_id) %>%
  arrange (icu_admittime) %>% 
  filter(row_number()==1) %>%
  select(apsiii:renal) %>%
  arrange(subject_id)

#Merging two grouped datasets
final_dataset_hadmid <- group_dataset_1 %>%
  left_join(group_dataset_2, by = c("subject_id","hadm_id"))

# Select subject_id from first hadm_id 
final_dataset <- final_dataset_hadmid %>%
  group_by(subject_id) %>%
  arrange (icu_admittime) %>% 
  filter(row_number() == 1) %>%
  arrange(subject_id)

```
# Export file

```{r}
write.csv(final_dataset,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/final_dataset.csv', row.names = F)
```

