---
title: "01_dataset_creation"
author: "Álvaro Ritoré"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
# output:
#   html_notebook:
#     code_folding: hide
#     number_sections: yes
#     theme: flatly
#     toc: yes
#     toc_float: yes
---

# Environment

```{r message=FALSE, warning=FALSE}
library(bigrquery)
library(DBI)
library(dplyr)
library(summarytools)
library(sqldf)
library(tidyverse)
library(nlme)
library(zoo)
library(tableone)
library(readxl)
library(feather)
library(magrittr)
library(MatchIt)
library(kableExtra)
library(ggplot2)
library(naniar)
library(modeest)
```

## BigQuery related functions

This chunks creates the run_query and getSQL function.

```{r}

# Function that takes in a SQL command and runs it on BigQuery. This avoids the connection data in each iteration
run_query<-function(query){
  query_output<-dbGetQuery(con,query)
  return(query_output)
}

# Function for reading SQL files from a folder
getSQL <- function(filepath) {
  con = file(filepath, "r")
  sql.string <- ""
  while (TRUE) {
    line <- readLines(con, n = 1)
    if (length(line) == 0) {
      break
    }
    line <- gsub("\\t", " ", line)
    if (grepl("--", line) == TRUE) {
      line <- paste(sub("--", "/*", line), "*/")
    }
    sql.string <- paste(sql.string, line)
  }
  close(con)
  return(sql.string)
}
```

## Setting up connection with BigQuery

```{r}
bigrquery::bq_auth() # UNCOMMENT WHEN HAVEN TOCKEN ISSUES!!!
# Establecer la conexión a BigQuery
project_id <- readline(prompt = "Enter your project ID: ")

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_id
)

```

# Data Load

Loading queries and extracting the data

## Prolonged Mechanical Ventilation

```{r}
# Criterion of prolonged mechanical ventilation: highest number of consecutive days of mechanical ventilation six or more hours per day
vm_consecutive_days_eicu <- run_query(getSQL("sql/vm_consecutive_days_eicu.sql"))

# Criterion of prolonged mechanical ventilation: total consecutive or non-consecutive days of mechanical ventilation for six or more hours per day
vm_total_days_eicu <- run_query(getSQL("sql/vm_total_days_eicu.sql"))

# Criterion of prolonged mechanical ventilation: 96 hour ventilation + tracheotomy
vm_and_traq_eicu <- run_query(getSQL("sql/vm_and_traq_eicu.sql"))

```

## Traumatisms

```{r}
#Traumatic Brain Injury
tbi_eicu <- run_query(getSQL("sql/tbi_eicu.sql"))
tbi_eicu %<>%
  group_by(stay_id) %>%
  summarize(
    tbi_seq1=max(tbi_min_diagnosisoffset), 
    tbi_anyseq=max(tbi_any_diagnosisoffset))

#General traumatisms
trauma_general_eicu <- run_query(getSQL("sql/trauma_general_eicu.sql"))
trauma_general_eicu %<>%
  group_by(stay_id) %>%
  summarize(
    traumatic_general_min_diagnosisoffset=max(traumatic_general_min_diagnosisoffset), 
    traumatic_general_any_diagnosisoffset=max(traumatic_general_any_diagnosisoffset))

#Fractures
trauma_fractures_eicu <- run_query(getSQL("sql/trauma_fractures_eicu.sql"))
trauma_fractures_eicu %<>%
  group_by(stay_id) %>%
  summarize(
    fractures_min_diagnosisoffset=max(fractures_min_diagnosisoffset),
    fractures_any_diagnosisoffset=max(fractures_any_diagnosisoffset))

#Luxations
trauma_luxations_eicu <- run_query(getSQL("sql/trauma_luxations_eicu.sql"))
trauma_luxations_eicu %<>%
  group_by(stay_id) %>%
  summarize(
    luxation_min_diagnosisoffset=max(luxation_min_diagnosisoffset),
    luxation_any_diagnosisoffset=max(luxation_any_diagnosisoffset))

#Sprains
trauma_sprains_eicu <- run_query(getSQL("sql/trauma_sprains_eicu.sql"))

#Burns
trauma_burns_eicu <- run_query(getSQL("sql/trauma_burns_eicu.sql"))
trauma_burns_eicu %<>%
  group_by(stay_id) %>%
  summarize(
    burns_min_diagnosisoffset=max(burns_min_diagnosisoffset),
    burns_any_diagnosisoffset=max(burns_any_diagnosisoffset))

#Contusion
trauma_contusion_eicu <- run_query(getSQL("sql/trauma_contusions_eicu.sql"))

```

## Demographic information

```{r}
#Patients
patient_eicu <- run_query(getSQL("sql/patient_eicu.sql"))

#Stays in ICU
icustays_eicu <- run_query(getSQL("sql/icustays_eicu.sql"))
```

## Severity scores

```{r}
#Table apachepatientresult 
apachepatientresult_eicu <- run_query(getSQL("sql/apachepatientresult_eicu.sql"))
apachepatientresult_eicu %<>%
  group_by(patientunitstayid) %>%
  arrange(apachepatientresultsid) %>%
  filter(row_number() == 1)

#Table apachepredvar 
apachepredvar_eicu <- run_query(getSQL("sql/apachepredvar_eicu.sql"))

#Table charlson 
charlson_eicu <- run_query(getSQL("sql/charlson_eicu.sql"))
```

# Creation of Final Dataset

## Joining across datasets

```{r}
# Join tables by patientunitstayid, departing from eICU's patient table
dataset_join <- Reduce(function(...) merge(..., by = c("patientunitstayid"), all.x=TRUE, sort = TRUE), list(
  patient_eicu,
  icustays_eicu, 
  apachepatientresult_eicu, 
  apachepredvar_eicu, 
  charlson_eicu))

# Rename patientunitstayid to stay_id
dataset_join %<>%
  rename("stay_id" = patientunitstayid)

# Join tables by stay_id
dataset_join <- Reduce(function(...) merge(..., by = c("stay_id"), all.x=TRUE, sort = TRUE), list(
  dataset_join, 
  vm_consecutive_days_eicu, 
  vm_total_days_eicu,
  vm_and_traq_eicu, 
  trauma_burns_eicu, 
  tbi_eicu,
  trauma_contusion_eicu, 
  trauma_fractures_eicu, 
  trauma_general_eicu, 
  trauma_luxations_eicu, 
  trauma_sprains_eicu))

```

## Creation of new variables and adjustment

```{r}
# Convert columns names to lowercase
names(dataset_join) <- tolower(names(dataset_join))

# Rename columns for self-explanation purposes
total_dataset <- dataset_join %>%
  rename(
    "subject_id_eicu" = uniquepid.x, #individual code for each patient
    "hadm_id_eicu" = patienthealthsystemstayid.x, #individual code for each hospital admission
    "age_eicu" = age.x, #age of the patient in full years. If the patient is over 89 years old specify “> 89” e.g.: 79, 36, 52, “> 89”, etc.
    "unitvisitnumber_eicu" = unitvisitnumber.x, #identifies the visit number of the patient, i.e. if the patient’s hospital stay has multiple unit stays
    "unitdischargeoffset" = unitdischargeoffset.x, #number of minutes from unit admit time that the patient was discharged from the unit
    "inhospital_death" = hospitaldischargestatus, #patient death within the given hospital stay from patient table
    "inicu_death" = unitdischargestatus, #patient death within the given ICU stay from patient table
    "discharge_location" = hospitaldischargelocation, #location where the patient was discharged to from the hospital e.g.: Home, Nursing Home, Death, etc.
    "aps" = acutephysiologyscore, #APS score (IV or IVa versions) from apachepatientresults table
    "apache" = apachescore, #Apache Score. Calculated from acutePhysiologyScore
    "charlson" = final_charlson_score, #comorbidity score charlson from SQL code calculating charlson index manually
    "gcs_verbal" = day1verbal, #from apachepredvar table
    "gcs_motor" = day1motor, #from apachepredvar table
    "gcs_eyes" = day1eyes #from apachepredvar table
    )

#Transform variables
total_dataset %<>%
  mutate(age_eicu = ifelse(age_eicu == "", NA, age_eicu), #Convert blank values to NA
         age_eicu = as.integer(ifelse(age_eicu == "> 89", 89, age_eicu)), #Convert >89 age values to 89
         sex = ifelse(gender.x!="Female" & gender.x!="Male", "Unknown", gender.x), #converts unknown sex values to Unknown
         hospital_stay_days = round((hospitaldischargeoffset.x - hospitaladmitoffset.x)/1440,2), #Convert minutes to days
         icu_stay_days = round(unitdischargeoffset/1440, 2), #Convert minutes to days
         discharge_location = ifelse(discharge_location == "", NA, discharge_location), #Convert blank values to NA
         inhospital_death = case_when(inhospital_death == "Expired" ~ 1, inhospital_death == "Alive" ~ 0, inhospital_death == "" ~ NA), #Convert Expired to 1, Alive to 0 and null to NA
         inicu_death = case_when(inicu_death == "Expired" ~ 1, inicu_death == "Alive" ~ 0, inicu_death == "" ~ NA), #Convert Expired to 1, Alive to 0 and null to NA
         inhospital_death = ifelse(is.na(inhospital_death) & inicu_death == 1, 1, inhospital_death), #for NA values in inhospital_death variable, fill with 1 when the patient died in ICU, since if the patient died in ICU therefore he died in hospital as well
         discharge_location = ifelse(is.na(discharge_location) & inhospital_death == 1, "Death", discharge_location) #for NA values in discharge_location variable, fill with "Death" when the patient died in ICU
          )

# Create variable days_of_death and death_30days
total_dataset %<>%
  mutate(days_of_death = as.numeric(ifelse(discharge_location == "Death" | is.na(discharge_location), hospital_stay_days, -1)), #if the patient is dead, the hospital length of stay represents the days of death from hospital admission. If the patient is alive, a value of -1 is given  
         death_30days = ifelse(days_of_death <=30 & days_of_death != -1, 1, 0) #patient death before 30 days from ICU admission
         )

# Traumatism variables
total_dataset %<>%
  mutate(
    anytrauma_seq1 = as.integer(ifelse(tbi_seq1 == 1 | fractures_min_diagnosisoffset == 1 | traumatic_general_min_diagnosisoffset == 1 | luxation_min_diagnosisoffset == 1, 1, 0)),          
    anytrauma_anyseq = as.integer(ifelse(tbi_anyseq == 1 | fractures_any_diagnosisoffset == 1 | traumatic_general_any_diagnosisoffset == 1 | luxation_any_diagnosisoffset == 1, 1, 0)),
            )

# Variables mv_14days (prolonged mechanical ventilation, criterion of 14 days) and mv_96traq (prolonged mechanical ventilation, criterion of 96h of ventilation + traqueostomy)
total_dataset %<>%
  mutate(
    mv_consecutive_7days = as.integer(ifelse(consecutive_vm_days>=7, 1, 0)),
    mv_consecutive_14days = as.integer(ifelse(consecutive_vm_days>=14, 1, 0)),
    mv_consecutive_21days = as.integer(ifelse(consecutive_vm_days>=21, 1, 0)),
    mv_total_7days = as.integer(ifelse(total_vm_days>=7, 1, 0)),
    mv_total_14days = as.integer(ifelse(total_vm_days>=14, 1, 0)),
    mv_total_21days = as.integer(ifelse(total_vm_days>=21, 1, 0)),
    mv_96traq = as.integer(ifelse(airwaytype=="Tracheostomy" & consecutive_vm_days>=4, 1, 0)),
    prolonged_mec_vent = as.integer(ifelse((mv_consecutive_14days == 1 |
                                             mv_96traq == 1), 1, 0))
      )

```

## Selection of required columns

```{r}
# Select the desired columns and order the result
total_dataset_selected <- total_dataset %>%
  select(subject_id_eicu, hadm_id_eicu, stay_id, unitvisitnumber_eicu, age_eicu, sex, hospitaladmitoffset.x, unitdischargeoffset, hospitaldischargeoffset.x, hospital_stay_days, icu_stay_days, discharge_location, inhospital_death, inicu_death, days_of_death, death_30days, tbi_seq1, tbi_anyseq, anytrauma_seq1, anytrauma_anyseq, consecutive_vm_days, mv_consecutive_7days, mv_consecutive_14days, mv_consecutive_21days, total_vm_days, mv_total_7days, mv_total_14days, mv_total_21days, mv_96traq, prolonged_mec_vent, aps, apache, charlson, gcs_verbal, gcs_motor, gcs_eyes) %>%
  rename("subject_id" = subject_id_eicu,
         "hadm_id" = hadm_id_eicu,
         "age" = age_eicu,
         "hospitaladmitoffset" = hospitaladmitoffset.x, 
         "hospitaldischargeoffset" = hospitaldischargeoffset.x,
         "unitvisitnumber" = unitvisitnumber_eicu) %>% 
  arrange(subject_id, unitvisitnumber)
```

## Addressing missingness

```{r}
# Map missingness in total_dataset before NA filling  
##Columns with na values
print('Columns in total_dataset having NAs')
names(total_dataset_selected)[colSums(is.na(total_dataset_selected)) > 0]

print('Nº of columns with na values')
n_var_miss(total_dataset_selected)

##Plot columns and show descriptive table with NA values
miss_summary_preclean <- miss_var_summary(total_dataset_selected)
miss_summary_preclean
gg_miss_var(total_dataset_selected, show_pct=TRUE)
  
```
## NA filling
```{r}
# Columns NA filling
total_dataset_clean <- total_dataset_selected %>%
  mutate(
    age = ifelse(is.na(age), modeest::mfv(age, na_rm=TRUE), age), #NAs are filled with the most frequent value (mode)
    sex = ifelse(is.na(sex), "Unknown", sex), #NA values of sex are converted to "Unknown"
    discharge_location = as.factor(ifelse(is.na(discharge_location), modeest::mfv(discharge_location), discharge_location)), # as a non-dichotomous categorical variable, NAs are filled with the most frequent value (mode)
    across(c(inhospital_death, inicu_death, death_30days:prolonged_mec_vent), ~ifelse(is.na(.), 0, .)), #NA values from dichotomous variables are filled with 0
    across(c(inhospital_death, inicu_death, death_30days:prolonged_mec_vent), as.integer),
    #apsiv and apache variables are omitted since they display a 26% of NA values
    aps = as.integer(ifelse(is.na(aps), modeest::mfv(aps, na_rm=TRUE), aps)), #NAs are filled with the most frequent value (mode)
    apache = ifelse(is.na(apache), modeest::mfv(age, na_rm=TRUE), age), #NAs are filled with the most frequent value (mode)
    across(gcs_verbal:gcs_eyes, ~ifelse(. == -1, 1, .)), #convert values equal to -1 to 1, the minimal value for each gcs variable. This could be due in eICU database to various reasons, such as the patient being intubated, sedated, paralyzed or with an eye injury that prevents evaluating their response. In these cases, a value of -1 is assigned to indicate that the data is missing or not applicable.
    gcs_original = as.integer(gcs_verbal + gcs_motor + gcs_eyes), #original gcs value before NA imputation
    across(gcs_verbal:gcs_eyes, ~ifelse(is.na(.), modeest::mfv(., na_rm=TRUE), .)), #fill NA values with the mode
    gcs = as.integer(gcs_verbal + gcs_motor + gcs_eyes) #total gcs value is the sum of verbal, motor and eyes (minimum: 3, maximum: 15)
  )

```

## Confirm 0% missingness
```{r}
# Map missingness in total_dataset_clean after NA filling  
##Columns with na values
print('Columns in total_dataset_clean having NAs')
names(total_dataset_clean)[colSums(is.na(total_dataset_clean)) > 0]

print('Nº of columns with na values')
n_var_miss(total_dataset_clean)

##Plot columns and show descriptive table with NA values
miss_summary_postclean <- miss_var_summary(total_dataset_clean)
miss_summary_postclean
gg_miss_var(total_dataset_clean, show_pct=TRUE)
  
```

## Creation of additional variables
```{r}
# Create variable tbi severe and non-severe tbi, as well as trauma_type, containing 4 types of trauma in our dataset: TBI, general traumatism without TBI, severe TBI, non-severe TBI, for both seq1 and anyseq 
total_dataset_clean %<>%
  mutate(
    tbi_severe_seq1 = as.integer(ifelse(tbi_seq1 == 1 & gcs <= 8, 1, 0)),
    tbi_nonsevere_seq1 = as.integer(ifelse(tbi_seq1 == 1 & gcs > 8, 1, 0)),
    trauma_type_seq1 = as.factor(
      ifelse(anytrauma_seq1 == 0, "non_trauma",
             ifelse(anytrauma_seq1 == 1 & tbi_seq1 == 0, "trauma_non_tbi",
                    ifelse(tbi_severe_seq1 == 1, "tbi_severe",
                           ifelse(tbi_nonsevere_seq1 == 1, "tbi_non_severe", NA)))
      )
    ),
    tbi_severe_anyseq = as.integer(ifelse(tbi_anyseq == 1 & gcs <= 8, 1, 0)),
    tbi_nonsevere_anyseq = as.integer(ifelse(tbi_anyseq == 1 & gcs > 8, 1, 0)),
    trauma_type_anyseq = as.factor(
      ifelse(anytrauma_anyseq == 0, "non_trauma",
             ifelse(anytrauma_anyseq == 1 & tbi_anyseq == 0, "trauma_non_tbi",
                    ifelse(tbi_severe_anyseq == 1, "tbi_severe",
                           ifelse(tbi_nonsevere_anyseq == 1, "tbi_non_severe", NA)))
      )
    )
  )

# Convert categorical variables to factor
total_dataset_clean %<>%
  mutate(sex = as.factor(sex),
         sex = relevel(sex, ref = "Female"), #change reference level of factor sex variable to "Male"
         discharge_location = as.factor(discharge_location), 
         trauma_type_seq1 = as.factor(trauma_type_seq1), 
         trauma_type_seq1 = relevel(trauma_type_seq1, ref = "non_trauma"),
         trauma_type_anyseq = as.factor(trauma_type_anyseq), 
         trauma_type_anyseq = relevel(trauma_type_anyseq, ref = "non_trauma"))


# Select and order columns
total_dataset_clean_selected <- total_dataset_clean %>% 
  select(subject_id:anytrauma_anyseq, tbi_severe_seq1, tbi_nonsevere_seq1, trauma_type_seq1, tbi_severe_anyseq, tbi_nonsevere_anyseq, trauma_type_anyseq, consecutive_vm_days:prolonged_mec_vent, aps, apache, charlson, gcs_original, gcs, gcs_verbal, gcs_motor, gcs_eyes) %>%
  arrange(subject_id, unitvisitnumber)
```

## Final Dataset
```{r}
# Group all stay_id within joint hospital admissions

#Summarizing variables into hadm_id groups. The first ICU stay has been included, although the whole ICU stay for that hospital admission has been summarized
group_dataset_1 <- total_dataset_clean_selected %>%
  group_by(subject_id, hadm_id) %>%
  arrange (unitvisitnumber) %>% 
  summarize(
    age=min(age), 
    sex=first(sex), 
    unitvisitnumber=first(unitvisitnumber), 
    hospitaladmitoffset=first(hospitaladmitoffset),
    unitdischargeoffset=last(unitdischargeoffset),
    hospitaldischargeoffset=first(hospitaldischargeoffset),
    hospital_stay_days=first(hospital_stay_days),
    icu_stay_days=sum(icu_stay_days), 
    discharge_location=last(discharge_location),
    inhospital_death=max(inhospital_death), 
    inicu_death=max(inicu_death), 
    days_of_death = max(days_of_death),
    death_30days=max(death_30days), 
    tbi_seq1=first(tbi_seq1), 
    tbi_anyseq=first(tbi_anyseq), 
    anytrauma_seq1=first(anytrauma_seq1),
    anytrauma_anyseq=first(anytrauma_anyseq),
    tbi_severe_seq1=first(tbi_severe_seq1), 
    tbi_nonsevere_seq1=first(tbi_nonsevere_seq1),
    trauma_type_seq1=first(trauma_type_seq1),
    tbi_severe_anyseq=first(tbi_severe_anyseq), 
    tbi_nonsevere_anyseq=first(tbi_nonsevere_anyseq),
    trauma_type_anyseq=first(trauma_type_anyseq),
    consecutive_vm_days=max(consecutive_vm_days), 
    mv_consecutive_7days=max(mv_consecutive_7days), 
    mv_consecutive_14days=max(mv_consecutive_14days),
    mv_consecutive_21days=max(mv_consecutive_21days),
    total_vm_days=sum(total_vm_days), 
    mv_total_7days = ifelse(total_vm_days>=7, 1, 0),
    mv_total_14days = ifelse(total_vm_days>=14, 1, 0),
    mv_total_21days = ifelse(total_vm_days>=21, 1, 0),
    mv_96traq=max(mv_96traq),
    prolonged_mec_vent=max(prolonged_mec_vent), 
          ) %>%
    arrange(subject_id)

#Select severity and comorbidity scores related to first ICU stay for each hospital admission   
group_dataset_2 <- total_dataset_clean_selected %>%
  group_by(subject_id, hadm_id) %>%
  arrange (unitvisitnumber) %>% 
  filter(row_number()==1) %>%
  select(aps:gcs_eyes) %>%
  arrange(subject_id)

#Merging two grouped datasets
final_dataset_hadmid <- group_dataset_1 %>%
  left_join(group_dataset_2, by = c("subject_id","hadm_id"))

#Print flowchart grouping criteria
cat(sprintf('%d unique patients in ICU in original eICU database.\n', nrow(total_dataset_clean_selected)))
rows_excluded_hadmid <- nrow(total_dataset_clean_selected)-nrow(group_dataset_1)
cat(sprintf('%d patients with unique hospital id, excluding %d rows.\n', nrow(group_dataset_1), rows_excluded_hadmid))


```

## Exclusion criteria
```{r}
# Apply exclusion filters and print rows excluded

## Removing under age patients
final_dataset_hadmid_excluded1 <- final_dataset_hadmid %>%
  filter(age>=18)
rows_excluded_age <- nrow(final_dataset_hadmid)-nrow(final_dataset_hadmid_excluded1)
cat(sprintf('%d patients remaining, with %d patients having <18 years old.\n', nrow(final_dataset_hadmid_excluded1), rows_excluded_age))

## Removing negative hospital admission offset
final_dataset_hadmid_excluded2 <- final_dataset_hadmid_excluded1 %>%
  filter(hospitaladmitoffset<=0) #the value should be negative as it is "number of minutes from unit admit time that the patient was admitted to the hospital", and otherwise it would mean ICU admission happened before hospital admission
rows_excluded_hospadmit <- nrow(final_dataset_hadmid_excluded1)-nrow(final_dataset_hadmid_excluded2)
cat(sprintf('%d patients remaining, with %d patients having hospital admission prior to ICU admission.\n', nrow(final_dataset_hadmid_excluded2), rows_excluded_hospadmit))

## Removing negative ICU length of stay
final_dataset_hadmid_excluded3 <- final_dataset_hadmid_excluded2 %>%
  filter(icu_stay_days>=0) 
rows_excluded_los <- nrow(final_dataset_hadmid_excluded2)-nrow(final_dataset_hadmid_excluded3)
cat(sprintf('%d patients remaining, with %d patients having negative ICU length of stay.\n', nrow(final_dataset_hadmid_excluded3), rows_excluded_los))

```


# Export file

```{r}
write.csv(final_dataset_hadmid_excluded,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/eICU/final_dataset_eICU.csv', row.names = F)
```

# Load File

```{r}
# Load file
final_dataset_eICU <- read.csv('C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/eICU/final_dataset_eICU.csv')

# Print final dataset and data frame summary
View(final_dataset_eICU)
stview(dfSummary(final_dataset_eICU))
```
