---
title: "03_modeling"
# output: html_document
---

# Environment

```{r}
# Libraries
library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(flextable)
library(officer)
library(rlang)
library(randomForest)
library(caret)
library(kableExtra)
library(magick)
library(gridExtra)
library(webshot2)

# Function and variable creation
## Define the root directory
root_dir <- "C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent"

## Function for recategorizing column to a specific variable
recategorize_function <- function(df, df_column, column_value) {
  df %>%
    mutate(!!df_column := as.factor(.data[[df_column]]),
           !!df_column := relevel(as.factor(.data[[df_column]]), ref = column_value))
}

## Function for extracting OR
extract_coef_pval_OR <- function(model) {
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_upper = round(exp(estimate + 1.96 * std.error),2)))
}

## Function for transforming names from OR's table and add adverse_protective variable
or_transformation <- function(df, custom_terms, variable_index) {
  df_transformed <- df %>%
    mutate(
      term = case_when(
        term == "(Intercept)" ~ "Intercept",
        TRUE ~ coalesce(custom_terms[term], term)  # Custom terms mapping or keep original
      ),
      adverse_protective = case_when(
        OR_lower > 1 & OR_upper > 1 ~ "Positive Association",
        OR_lower < 1 & OR_upper < 1 ~ "Negative Association",
        TRUE ~ "Inconclusive"
      ),
      adverse_protective = factor(
        adverse_protective,
        levels = c("Positive Association", "Negative Association", "Inconclusive")
      )
    ) %>%
    arrange(match(term, variable_index)) %>%
    filter(term != "Intercept")

  rownames(df) <- NULL #resets index

  return(df_transformed)
}


## Dictionary to relabel terms in the final tables (for better readability and presentation)
dict_terms <- c(
  "prolonged_mec_vent" = "PMV",
  "trauma_type_anyseqnon-severe TBI" = "TBI non-severe",
  "trauma_type_anyseqsevere TBI" = "TBI severe",
  "trauma_type_anyseqnon-TBI trauma" = "non-TBI trauma",
  "trauma_type_anyseqnon-trauma" = "non-trauma",
  "prolonged_mec_vent:trauma_type_anyseqsevere TBI" = "PMV + TBI severe",
  "prolonged_mec_vent:trauma_type_anyseqnon-severe TBI" = "PMV + TBI non-severe",
  "prolonged_mec_vent:trauma_type_anyseqnon-TBI trauma" = "PMV + non-TBI trauma",
  "prolonged_mec_vent:trauma_type_anyseqnon-trauma" = "PMV + non-trauma",
  "trauma_type3_anyseqnon-TBI trauma" = "non-TBI trauma",
  "trauma_type3_anyseqnon-trauma" = "non-trauma",
  "trauma_type3_anyseqTBI" = "TBI",
  "prolonged_mec_vent:trauma_type3_anyseqnon-TBI trauma" = "PMV + non-TBI trauma",
  "prolonged_mec_vent:trauma_type3_anyseqnon-trauma" = "PMV + non-trauma",
  "prolonged_mec_vent:trauma_type3_anyseqTBI" = "PMV + TBI",
  "anytrauma_anyseq0" = "non-trauma",
  "prolonged_mec_vent:anytrauma_anyseq0" = "PMV + non-trauma",
  "anytrauma_anyseq1" = "Any Trauma",
  "prolonged_mec_vent:anytrauma_anyseq1" = "PMV + Any Trauma",
  "age" = "Age (years)",
  "sexMale" = "Sex (male)",
  "sexUnknown" = "Sex (unknown)",
  "aps" = "APSIII",
  "charlson" = "Charlson",
  "PMV for non-Trauma" = "PMV - non-trauma",
  "PMV for Trauma not TBI" = "PMV - non-TBI trauma",
  "PMV for TBI non-severe" = "PMV - non-severe TBI",
  "PMV for TBI severe" = "PMV - severe TBI",
  "metastasis" = "Metastasis",
  "hiv" = "HIV",
  "liver_cirrhosis" = "Liver Cirrhosis",
  "stroke" = "Stroke",
  "renal_disease" = "Renal Disease",
  "diabetes_disease" = "Diabetes",
  "cancer" = "Cancer",
  "leukemia_disease" = "Leukemia",
  "lymphoma_disease" = "Lymphoma",
  "myocardial_infarction" = "Myocardial Infarction",
  "chf" = "Congestive Heart Failure (CHF)",
  "pvd" = "Peripheral Vascular Disease (PVD)",
  "tia" = "Transient Ischemic Attack (TIA)",
  "dementia" = "Dementia",
  "copd" = "Chronic Obstructive Pulmonary Disease (COPD)",
  "ctd" = "Connective Tissue Disease (CTD)",
  "pud" = "Peptic Ulcer Disease (PUD)",
  "liver_disease" = "Liver Disease"
)

## Variable order in OR table
variable_order <- c(
  "Intercept", "PMV", "Age (years)", "Sex (male)", "Sex (unknown)", "APSIII", "Charlson", "non-trauma", "non-TBI trauma", "TBI", "TBI non-severe", "TBI severe", "PMV + non-trauma", "PMV + non-TBI trauma", "PMV + TBI", "PMV + TBI non-severe", "PMV + TBI severe")

## Variables for table output in console
var_html <- c("term", "p.value", "OR","OR_lower","OR_upper")

## Function for performing univariate analysis
univariate_analysis <- function(df, death_outcome, df_column=NULL, output_name) {
  ## If df_column is not specified, it takes the whole dataset
  if (is.null(df_column)) {
    model <- glm(df[[death_outcome]] ~ prolonged_mec_vent,
                 data = df,
                 family = binomial())
  } else {
    model <- glm(df[[death_outcome]] ~ prolonged_mec_vent * df[[df_column]],
                 data = df,
                 family = binomial())
  }
  or_univariate <- extract_coef_pval_OR(model)
  row_name <- or_univariate[or_univariate[, 1] == "prolonged_mec_vent", var_html]
  row_name[, 1] <- output_name
  html_table_output <- htmlTable(row_name[,var_html], caption = paste("Univariate Analysis. Odds Ratio and Confidence Intervals for factors associated with", death_outcome, "- eICU Database"))
  print(html_table_output)
  return(row_name)
}

## Function to obtain counts and mortality percentages for different columns and categories
create_summary_table <- function(df, death_outcome, df_column=NULL, row_label, column_value=NULL) {

  ## If df_column is not specified, it takes the whole dataset. If it is, a subset is taken
  if (is.null(df_column)) {
    subset_trauma <- df
  } else {
    subset_trauma <- subset(df, df[[df_column]] == column_value)
  }

  ## Filters dataframe for specific column and trauma type
  subset_trauma_pmv <- subset(subset_trauma, prolonged_mec_vent == 1)
  subset_trauma_nonpmv <- subset(subset_trauma, prolonged_mec_vent == 0)

  ## Calculate totals and percentages
  total_number <- nrow(subset_trauma)
  Pct_category <- (total_number / nrow(df)) * 100
  n_PMV <- nrow(subset_trauma_pmv)
  n_not_pmv <- nrow(subset_trauma_nonpmv)

  Pct_PMV <- (n_PMV / total_number) * 100
  Pct_not_PMV <- (n_not_pmv / total_number) * 100

  ## Create summary table
  summary_table <- data.frame(
    trauma_type = row_label,
    total_number = total_number,
    Pct_category = Pct_category,
    n_PMV = n_PMV,
    Pct_PMV = Pct_PMV,
    mortality_pmv = (sum(subset_trauma_pmv[[death_outcome]] == 1) / n_PMV) * 100,
    n_not_pmv = n_not_pmv,
    Pct_not_PMV = Pct_not_PMV,
    mortality_not_pmv = (sum(subset_trauma_nonpmv[[death_outcome]] == 1) / n_not_pmv) * 100
  )

  return(summary_table)
}

## Function for multivariable mortality analysis with GAM age spline (sensitivity)
run_age_spline_model <- function(df,
                                 model_formula,
                                 death_outcome,
                                 output_name,
                                 output_file,
                                 k = 10) {

  cat(sprintf("\nLogistic Regression of %s: %s + age(spline)\n", death_outcome, output_name))

  if (!exists("root_dir", inherits = TRUE)) {
    stop("root_dir is not defined. Define root_dir before running spline model exports.")
  }

  # Replace the parametric 'age' term by a thin-plate spline s(age).
  # k is only the basis dimension (upper bound on flexibility), not the final edf.
  gam_formula <- update(
    model_formula,
    . ~ . - age + s(age, k = k, bs = "tp")
  )
  # Reset the formula environment so that the local argument 'k' is resolved by gam().
  environment(gam_formula) <- environment()

  # Spline model: smoothness (edf) is estimated automatically via the REML penalty.
  logreg_age_spline <- mgcv::gam(
    gam_formula,
    data = df,
    family = binomial(link = "logit"),
    method = "REML"
  )

  # Reference linear-age model (same engine) used as the linearity comparator.
  logreg_age_linear <- mgcv::gam(
    model_formula,
    data = df,
    family = binomial(link = "logit"),
    method = "REML"
  )

  gam_summary <- summary(logreg_age_spline)
  age_smooth_row <- gam_summary$s.table["s(age)", ]

  # Smooth summary: edf near 1 suggests a linear age-outcome relationship.
  age_spline_summary <- data.frame(
    group = paste0(output_name, " + age(spline)"),
    outcome = death_outcome,
    smooth = sprintf("s(age, k = %s, bs = 'tp')", k),
    age_edf = age_smooth_row[["edf"]],
    age_smooth_pvalue = age_smooth_row[["p-value"]]
  )

  # OR table for parametric terms only (spline basis coefficients are removed).
  or_1_age_spline <- extract_coef_pval_OR(logreg_age_spline)
  or_1_age_spline <- or_1_age_spline[!grepl("^s\\(age\\)", or_1_age_spline$term), ]
  or_2_age_spline <- or_transformation(or_1_age_spline, dict_terms, variable_order)
  or_pmv_age_spline <- or_2_age_spline[or_2_age_spline[, 1] == "PMV", c(1, 5:9)]
  or_pmv_age_spline[, 1] <- paste0(output_name, " + age(spline)")

  output_path <- file.path(root_dir, output_file)
  output_smooth_path <- file.path(root_dir, sub("\\.csv$", "_smooth.csv", output_file))

  write.csv(or_2_age_spline, output_path)
  write.csv(
    age_spline_summary,
    output_smooth_path,
    row.names = FALSE
  )

  # k-adequacy diagnostic: if edf is close to k' with a low p-value, increase k.
  set.seed(42)
  k_diagnostic <- mgcv::k.check(logreg_age_spline)

  # Linearity diagnostic: AIC of the linear-age vs the spline-age model.
  aic_comparison <- AIC(logreg_age_linear, logreg_age_spline)

  # ---- Console / knit output ----
  print(htmlTable(or_2_age_spline[, var_html], caption = caption_hospdeath_spline_multivariable))

  cat("\n-- Age smooth summary (edf, smooth p-value) --\n")
  print(age_spline_summary)

  cat("\n-- k-adequacy check (k.check): edf should stay below k' --\n")
  print(k_diagnostic)

  cat("\n-- Linearity check: AIC (linear age vs spline age) --\n")
  print(aic_comparison)

  # Smooth on the logit scale: a near-straight line corroborates linearity.
  plot(logreg_age_spline,
       select = 1, shade = TRUE, seWithMean = TRUE,
       main = paste0(output_name, ": s(age) on logit scale"))

  return(list(
    model = logreg_age_spline,
    linear_model = logreg_age_linear,
    or_table = or_2_age_spline,
    or_pmv = or_pmv_age_spline,
    gam_summary = gam_summary,
    age_spline_summary = age_spline_summary,
    k_check = k_diagnostic,
    aic_comparison = aic_comparison
  ))
}

## Function to build and export the GAM sensitivity comparison table to .docx
## Inputs:
##   - linear_or_list:    list of one-row data.frames with the linear PMV OR per subgroup
##                        (each row has columns: term, p.value, OR, OR_lower, OR_upper, adverse_protective)
##   - spline_result_list: list of run_age_spline_model() returns per subgroup
##   - subgroup_labels:   character vector with the labels to display (in display order, top to bottom)
##   - caption:           table caption (string)
##   - output_dir:        relative path under root_dir where the .docx will be saved
##   - save_name:         filename (without extension) for the .docx
## The two lists must be in the SAME order as subgroup_labels.

build_age_spline_comparison_table <- function(linear_or_list,
                                              spline_result_list,
                                              subgroup_labels,
                                              caption,
                                              output_dir,
                                              save_name) {

  # Build one row per subgroup combining linear OR, spline OR, edf, and ΔAIC
  rows <- lapply(seq_along(subgroup_labels), function(i) {
    lin <- linear_or_list[[i]]
    spl <- spline_result_list[[i]]
    spl_or <- spl$or_pmv

    # Format OR (95% CI) as "X.XX (X.XX–X.XX)"
    or_lin_fmt <- sprintf("%.2f (%.2f–%.2f)", lin$OR, lin$OR_lower, lin$OR_upper)
    or_spl_fmt <- sprintf("%.2f (%.2f–%.2f)", spl_or$OR, spl_or$OR_lower, spl_or$OR_upper)

    # Use the same rounding as the OR tables to keep Δ consistent with reported OR
    delta_or  <- round(spl_or$OR, 2) - round(lin$OR, 2)
    edf_val   <- round(spl$age_spline_summary$age_edf, 2)
    delta_aic <- round(spl$aic_comparison$AIC[1] - spl$aic_comparison$AIC[2], 1)

    data.frame(
      Subgroup            = subgroup_labels[i],
      `OR (95% CI), linear age`  = or_lin_fmt,
      `OR (95% CI), spline age`  = or_spl_fmt,
      `Δ OR`              = sprintf("%+.2f", delta_or),
      `edf`               = edf_val,
      `ΔAIC`              = delta_aic,
      check.names = FALSE,
      stringsAsFactors = FALSE
    )
  })

  comparison_df <- do.call(rbind, rows)
  dir.create(file.path(root_dir, output_dir), recursive = TRUE, showWarnings = FALSE)

  # Build the flextable
  set_flextable_defaults(font.size = 10)
  ft <- flextable(comparison_df) %>%
    set_caption(caption = caption) %>%
    set_table_properties(layout = "autofit") %>%
    align(align = "left",   part = "body",   j = 1) %>%
    align(align = "center", part = "body",   j = 2:ncol(comparison_df)) %>%
    align(align = "center", part = "header") %>%
    bold(part = "header") %>%
    add_footer_lines(values = paste(
      "edf: effective degrees of freedom of the age smooth.",
      "ΔAIC = AIC(linear age) − AIC(spline age); positive values favor the spline.",
      "All GAMs fitted with thin-plate regression splines (k = 10, REML).",
      "Note: subgroups sharing the same trauma categorization (e.g., non-severe and severe TBI)",
      "are relevels of the same underlying GAM, so they share edf and ΔAIC."
    ))

  # Export to .docx
  read_docx() %>%
    body_add_flextable(value = ft) %>%
    print(target = file.path(root_dir, output_dir, paste0(save_name, ".docx")))

  return(comparison_df)
}


# Titles for html outputs
caption_hospdeath_multivariable <- "Multivariable Analysis. Odds Ratio and Confidence Intervals for Hospital Mortality-associated Factors. eICU Database"
caption_hospdeath_spline_multivariable <- "Multivariable Analysis with age(spline). Odds Ratio and Confidence Intervals for Hospital Mortality-associated Factors. eICU Database"

```

# Load File

```{r}
final_dataset_eICU <- readRDS("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/eICU/final_dataset_eICU.rds")
```

# Hospital Death

## Total Population

```{r}
# Multivariate analysis
## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent + age + sex + aps + charlson,
  data = final_dataset_eICU,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_allpopulation_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_allpopulation_eICU_multivariate[, 1] <- "PMV for All Population"

## Save OR table and print in console
write.csv(or_2, file.path(root_dir,'data/eICU/or_hospdeath_PMV-allpopulation_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_allpopulation_eICU_univariate <- univariate_analysis(df=final_dataset_eICU, death_outcome="inhospital_death", output_name="PMV for All Population")
```

## Total Population + age(spline)

```{r}
hospdeath_age_spline_allpopulation_eICU <- run_age_spline_model(
  df = final_dataset_eICU,
  model_formula = inhospital_death ~ prolonged_mec_vent + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for All Population",
  output_file = "data/eICU/or_hospdeath_PMV-allpopulation_age-spline_eICU.csv"
)

or_hospdeath_PMV_allpopulation_age_spline_eICU_multivariate <- hospdeath_age_spline_allpopulation_eICU$or_pmv
```

## non_trauma

```{r}
# Multivariate analysis
## Recategorize 0 (non-trauma) as reference level in anytrauma_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "anytrauma_anyseq", "0")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_nontrauma_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_nontrauma_eICU_multivariate[, 1] <- "PMV for non-Trauma"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-nontrauma_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_nontrauma_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="anytrauma_anyseq", output_name="PMV for non-Trauma")
```

## non_trauma + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "anytrauma_anyseq", "0")

hospdeath_age_spline_nontrauma_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for non-Trauma",
  output_file = "data/eICU/or_hospdeath_PMV-nontrauma_age-spline_eICU.csv"
)

or_hospdeath_PMV_nontrauma_age_spline_eICU_multivariate <- hospdeath_age_spline_nontrauma_eICU$or_pmv
```

## Traumatized

```{r}
# Multivariate analysis
## Recategorize 1 (trauma) as reference level in anytrauma_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "anytrauma_anyseq", "1")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_anytrauma_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_anytrauma_eICU_multivariate[, 1] <- "PMV for Traumatized Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-anytrauma_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_anytrauma_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="anytrauma_anyseq", output_name="PMV for Traumatized Patients")

```

## Traumatized + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "anytrauma_anyseq", "1")

hospdeath_age_spline_anytrauma_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for Traumatized Patients",
  output_file = "data/eICU/or_hospdeath_PMV-anytrauma_age-spline_eICU.csv"
)

or_hospdeath_PMV_anytrauma_age_spline_eICU_multivariate <- hospdeath_age_spline_anytrauma_eICU$or_pmv
```

## trauma_non_tbi

```{r}
# Multivariate Analysis
## Recategorize "non-TBI trauma" as reference level in trauma_type3_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type3_anyseq", "non-TBI trauma")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * trauma_type3_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_traumanontbi_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_traumanontbi_eICU_multivariate[, 1] <- "PMV for Trauma not TBI"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-TraumanotTBI_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_traumanottbi_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type3_anyseq", output_name="PMV for Trauma not TBI")
```

## trauma_non_tbi + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type3_anyseq", "non-TBI trauma")

hospdeath_age_spline_traumanontbi_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * trauma_type3_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for Trauma not TBI",
  output_file = "data/eICU/or_hospdeath_PMV-TraumanotTBI_age-spline_eICU.csv"
)

or_hospdeath_PMV_traumanontbi_age_spline_eICU_multivariate <- hospdeath_age_spline_traumanontbi_eICU$or_pmv
```

## TBI

```{r}
# Multivariate analysis
## Recategorize "TBI" as reference level in trauma_type3_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type3_anyseq", "TBI")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * trauma_type3_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbi_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbi_eICU_multivariate[, 1] <- "PMV for TBI Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-tbi_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_tbi_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type3_anyseq", output_name="PMV for TBI Patients")
```

## TBI + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type3_anyseq", "TBI")

hospdeath_age_spline_tbi_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * trauma_type3_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for TBI Patients",
  output_file = "data/eICU/or_hospdeath_PMV-tbi_age-spline_eICU.csv"
)

or_hospdeath_PMV_tbi_age_spline_eICU_multivariate <- hospdeath_age_spline_tbi_eICU$or_pmv
```

## tbi_non_severe

```{r}
# Multivariate Analysis
## Recategorize "non-severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type_anyseq", "non-severe TBI")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbinonsevere_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbinonsevere_eICU_multivariate[, 1] <- "PMV for TBI non-severe"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-TBInonsevere_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_nonseveretbi_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for TBI non-severe")

```

## tbi_non_severe + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type_anyseq", "non-severe TBI")

hospdeath_age_spline_tbinonsevere_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for TBI non-severe",
  output_file = "data/eICU/or_hospdeath_PMV-TBInonsevere_age-spline_eICU.csv"
)

or_hospdeath_PMV_tbinonsevere_age_spline_eICU_multivariate <- hospdeath_age_spline_tbinonsevere_eICU$or_pmv
```

## tbi_severe

```{r}
# Multivariate Analysis
## Recategorize "severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type_anyseq", "severe TBI")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
   inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + aps + charlson,
  data = final_dataset_recategorized,
  family = binomial())

## Extract OR, assign proper labels and save PMV's OR for this specific model
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbisevere_eICU_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbisevere_eICU_multivariate[, 1] <- "PMV for TBI severe"

# Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/eICU/or_hospdeath_PMV-TBIsevere_eICU.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_severetbi_eICU_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for TBI severe")
```

## tbi_severe + age(spline)

```{r}
final_dataset_recategorized <- recategorize_function(final_dataset_eICU, "trauma_type_anyseq", "severe TBI")

hospdeath_age_spline_tbisevere_eICU <- run_age_spline_model(
  df = final_dataset_recategorized,
  model_formula = inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + aps + charlson,
  death_outcome = "inhospital_death",
  output_name = "PMV for TBI severe",
  output_file = "data/eICU/or_hospdeath_PMV-TBIsevere_age-spline_eICU.csv"
)

or_hospdeath_PMV_tbisevere_age_spline_eICU_multivariate <- hospdeath_age_spline_tbisevere_eICU$or_pmv
```

## Summary

### Multivariate Analysis

```{r}
# Obtain combined table
eICU_or_hospdeath_PMV_multivariate <- rbind(
  or_hospdeath_PMV_allpopulation_eICU_multivariate,
  or_hospdeath_PMV_nontrauma_eICU_multivariate,
  or_hospdeath_PMV_anytrauma_eICU_multivariate,
  or_hospdeath_PMV_traumanontbi_eICU_multivariate,
  or_hospdeath_PMV_tbi_eICU_multivariate,
  or_hospdeath_PMV_tbinonsevere_eICU_multivariate,
  or_hospdeath_PMV_tbisevere_eICU_multivariate
) %>%
  mutate_if(is.numeric, round, 3)
rownames(eICU_or_hospdeath_PMV_multivariate) <- NULL
# Save OR table and print in console
write.csv(eICU_or_hospdeath_PMV_multivariate,file.path(root_dir,'data/eICU/or_hospdeath_PMV_all-trauma_eICU.csv'))
htmlTable(eICU_or_hospdeath_PMV_multivariate[,var_html], caption = "Multivariate Analysis. Odds Ratio and Confidence Intervals for factors associated with Hospital Mortality. All types of trauma. eICU Database")

# Assign proper names to OR table
eICU_or_hospdeath_PMV_multivariate_2 <- or_transformation(eICU_or_hospdeath_PMV_multivariate, dict_terms, variable_order)
# Order and select OR table for building Forest Plot
eICU_or_hospdeath_PMV_multivariate_3 <- eICU_or_hospdeath_PMV_multivariate_2 %>%
  filter(term %in% c("PMV - severe TBI", "PMV - non-severe TBI", "PMV - non-TBI trauma", "PMV - non-trauma")) %>%
  mutate(term = factor(term, levels = c("PMV - severe TBI", "PMV - non-severe TBI", "PMV - non-TBI trauma", "PMV - non-trauma"))) %>%
  arrange(term)

# Create a data frame with all possible levels for adverse-protective-inconclusive variable
adverse_protective_levels <- data.frame(adverse_protective = c("Positive Association", "Negative Association", "Inconclusive"))

# Build the Forest plot
fp_PMV_alltraumas_hospdeath_eICU <- ggplot(eICU_or_hospdeath_PMV_multivariate_3,
       aes(x = OR, xmin = OR_lower, xmax = OR_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "c. eICU. Outcome: Hospital Mortality") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Positive Association", "Negative Association", "Inconclusive"),
    name = "Effect on Outcome",
    limits = adverse_protective_levels$adverse_protective) + #establishes limits of color scale
  scale_x_continuous(limits = c(0, 6)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))
fp_PMV_alltraumas_hospdeath_eICU


# Save the Forest plot as a PNG file
ggsave(file.path(root_dir, 'figures/eICU/Forest_plot/fp_hospdeath_PMV_all-trauma_eICU.png'), plot=fp_PMV_alltraumas_hospdeath_eICU, width = 10, height = 4, dpi = 300)
```

### Multivariate Analysis + age(spline)

```{r}
# Obtain combined table
eICU_or_hospdeath_PMV_age_spline_multivariate <- rbind(
  or_hospdeath_PMV_allpopulation_age_spline_eICU_multivariate,
  or_hospdeath_PMV_nontrauma_age_spline_eICU_multivariate,
  or_hospdeath_PMV_anytrauma_age_spline_eICU_multivariate,
  or_hospdeath_PMV_traumanontbi_age_spline_eICU_multivariate,
  or_hospdeath_PMV_tbi_age_spline_eICU_multivariate,
  or_hospdeath_PMV_tbinonsevere_age_spline_eICU_multivariate,
  or_hospdeath_PMV_tbisevere_age_spline_eICU_multivariate
) %>%
  mutate_if(is.numeric, round, 3)
rownames(eICU_or_hospdeath_PMV_age_spline_multivariate) <- NULL

# Save OR table and print in console
write.csv(eICU_or_hospdeath_PMV_age_spline_multivariate,file.path(root_dir,'data/eICU/or_hospdeath_PMV_all-trauma_age-spline_eICU.csv'))
htmlTable(eICU_or_hospdeath_PMV_age_spline_multivariate[,var_html], caption = "Multivariable Analysis with age(spline). Odds Ratio and Confidence Intervals for factors associated with Hospital Mortality. All types of trauma. eICU Database")

# Order and select OR table for building Forest Plot
eICU_or_hospdeath_PMV_age_spline_multivariate_3 <- eICU_or_hospdeath_PMV_age_spline_multivariate %>%
  filter(term %in% c("PMV for TBI severe + age(spline)", "PMV for TBI non-severe + age(spline)", "PMV for Trauma not TBI + age(spline)", "PMV for non-Trauma + age(spline)")) %>%
  mutate(term = factor(term, levels = c("PMV for TBI severe + age(spline)", "PMV for TBI non-severe + age(spline)", "PMV for Trauma not TBI + age(spline)", "PMV for non-Trauma + age(spline)"))) %>%
  arrange(term)

# Build the Forest plot
fp_PMV_alltraumas_hospdeath_age_spline_eICU <- ggplot(eICU_or_hospdeath_PMV_age_spline_multivariate_3,
       aes(x = OR, xmin = OR_lower, xmax = OR_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "eICU. Outcome: Hospital Mortality + age(spline)") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Positive Association", "Negative Association", "Inconclusive"),
    name = "Effect on Outcome",
    limits = adverse_protective_levels$adverse_protective) + #establishes limits of color scale
  scale_x_continuous(limits = c(0, 6)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))
fp_PMV_alltraumas_hospdeath_age_spline_eICU


# Save the Forest plot as a PNG file
ggsave(file.path(root_dir, 'figures/eICU/Forest_plot/fp_hospdeath_PMV_all-trauma_age-spline_eICU.png'), plot=fp_PMV_alltraumas_hospdeath_age_spline_eICU, width = 10, height = 4, dpi = 300)

# Supplementary Table: GAM sensitivity comparison for in-hospital mortality (eICU).
hospdeath_eICU_spline_comparison <- build_age_spline_comparison_table(
  linear_or_list = list(
    or_hospdeath_PMV_nontrauma_eICU_multivariate,
    or_hospdeath_PMV_traumanontbi_eICU_multivariate,
    or_hospdeath_PMV_tbinonsevere_eICU_multivariate,
    or_hospdeath_PMV_tbisevere_eICU_multivariate
  ),
  spline_result_list = list(
    hospdeath_age_spline_nontrauma_eICU,
    hospdeath_age_spline_traumanontbi_eICU,
    hospdeath_age_spline_tbinonsevere_eICU,
    hospdeath_age_spline_tbisevere_eICU
  ),
  subgroup_labels = c("non-trauma", "non-TBI trauma", "non-severe TBI", "severe TBI"),
  caption = paste(
    "Supplementary Table Sz. Sensitivity analysis comparing PMV odds ratios for in-hospital",
    "mortality under linear vs. penalized spline modeling of age. eICU cohort."
  ),
  output_dir = "data/eICU/tables",
  save_name  = "suppl_hospdeath_PMV_spline-vs-linear_eICU"
)
```

### Univariate Analysis

```{r}
# Build univariate analysis table
hospdeath_univariate_summary <- rbind(
  or_hospdeath_PMV_allpopulation_eICU_univariate,
  or_hospdeath_PMV_nontrauma_eICU_univariate,
  or_hospdeath_PMV_anytrauma_eICU_univariate,
  or_hospdeath_PMV_traumanottbi_eICU_univariate,
  or_hospdeath_PMV_tbi_eICU_univariate,
  or_hospdeath_PMV_nonseveretbi_eICU_univariate,
  or_hospdeath_PMV_severetbi_eICU_univariate
) %>%
  mutate_if(is.numeric, round, 3)
rownames(hospdeath_univariate_summary) <- NULL
## Save OR table and print in console
write.csv(hospdeath_univariate_summary,file.path(root_dir,'data/eICU/hospdeath_univariate_summary_eICU.csv'))
htmlTable(hospdeath_univariate_summary, caption = "Univariate Analysis for different subgroups. eICU Database")
```

### Counts and Percentages

```{r}
# Obtain counts and mortality percentages for each trauma subset
all_summary <- create_summary_table(df=final_dataset_eICU, death_outcome="inhospital_death", row_label="All selected population")
any_trauma_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "anytrauma_anyseq", "Any Trauma", 1)
non_trauma_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "trauma_type_anyseq", "non-trauma", "non-trauma")
tbi_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "tbi_anyseq", "TBI", 1)
nontbi_trauma_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "trauma_type_anyseq", "non-TBI trauma", "non-TBI trauma")
nonsevere_tbi_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "trauma_type_anyseq", "non-severe TBI", "non-severe TBI")
severe_tbi_summary <- create_summary_table(final_dataset_eICU, death_outcome="inhospital_death", "trauma_type_anyseq", "severe TBI", "severe TBI")

## Build counts and percentages table
hospdeath_counts_summary <- rbind(
  all_summary,
  non_trauma_summary,
  any_trauma_summary,
  nontbi_trauma_summary,
  tbi_summary,
  nonsevere_tbi_summary,
  severe_tbi_summary
) %>%
  mutate_if(is.numeric, round, 2)
rownames(hospdeath_counts_summary) <- NULL
## Save OR table and print in console
write.csv(hospdeath_counts_summary,file.path(root_dir,'data/eICU/hospdeath_counts_summary_eICU.csv'))
htmlTable(hospdeath_counts_summary, caption = "Counts and percentages of different subgroups. eICU Database")
```


# Propensity Score Matching

## PSM
```{r}
# ============================================================
# PROPENSITY SCORE MATCHING (PSM) - SENSITIVITY ANALYSIS
#
# Strategy: within each trauma subgroup, match each PMV patient
# to the most similar non-PMV patient by propensity score, then
# re-estimate the OR of PMV on hospital mortality in the matched sample.
# ============================================================

library(MatchIt)
library(cobalt)
library(sandwich)
library(lmtest)

dir.create(file.path(root_dir, "figures/eICU/PSM"), recursive = TRUE, showWarnings = FALSE)

required_psm_vars <- c(
  "prolonged_mec_vent", "inhospital_death", "trauma_type_anyseq",
  "age", "sex", "aps", "charlson"
)

missing_psm_vars <- setdiff(required_psm_vars, names(final_dataset_eICU))
if (length(missing_psm_vars) > 0) {
  stop(sprintf("Missing variables required for eICU PSM: %s", paste(missing_psm_vars, collapse = ", ")))
}

run_psm_analysis_eicu <- function(df,
                                  death_outcome,
                                  group_label,
                                  covariates = c("age", "sex", "aps", "charlson"),
                                  ratio      = 1,
                                  caliper    = 0.2,
                                  root_dir) {

  missing_vars <- setdiff(c("prolonged_mec_vent", death_outcome, covariates), names(df))
  if (length(missing_vars) > 0) {
    stop(sprintf("Group '%s': missing variables for PSM: %s",
                 group_label, paste(missing_vars, collapse = ", ")))
  }

  cat(sprintf("\n========== eICU PSM: %s | Outcome: %s ==========\n",
              group_label, death_outcome))

  n_pmv   <- sum(df$prolonged_mec_vent == 1, na.rm = TRUE)
  n_nopmv <- sum(df$prolonged_mec_vent == 0, na.rm = TRUE)
  cat(sprintf("Pre-matching: PMV = %d, no-PMV = %d\n", n_pmv, n_nopmv))

  if (n_pmv < 10 || n_nopmv < 10) {
    warning(sprintf(
      "Group '%s': insufficient n for PSM (PMV=%d, no-PMV=%d). Skipping.",
      group_label, n_pmv, n_nopmv))
    return(NULL)
  }

  n_events_pmv   <- sum(df[[death_outcome]][df$prolonged_mec_vent == 1], na.rm = TRUE)
  n_events_nopmv <- sum(df[[death_outcome]][df$prolonged_mec_vent == 0], na.rm = TRUE)
  cat(sprintf("Events (pre-matching): PMV = %d, no-PMV = %d\n",
              n_events_pmv, n_events_nopmv))

  if (n_events_pmv < 5 || n_events_nopmv < 5) {
    warning(sprintf(
      "Group '%s': insufficient events for PSM (events PMV=%d, no-PMV=%d). Skipping.",
      group_label, n_events_pmv, n_events_nopmv))
    return(NULL)
  }

  ps_formula <- as.formula(
    paste("prolonged_mec_vent ~", paste(covariates, collapse = " + "))
  )

  match_obj <- matchit(
    formula     = ps_formula,
    data        = df,
    method      = "nearest",
    distance    = "glm",
    link        = "logit",
    replace     = FALSE,
    ratio       = ratio,
    caliper     = caliper,
    std.caliper = TRUE
  )

  matched_df  <- match.data(match_obj)
  n_matched   <- as.integer(sum(matched_df$prolonged_mec_vent == 1))
  n_unmatched <- as.integer(n_pmv - n_matched)
  cat(sprintf("Post-matching: %d pairs formed | PMV unmatched: %d\n",
              n_matched, n_unmatched))

  n_events_pmv_matched <- sum(
    matched_df[[death_outcome]][matched_df$prolonged_mec_vent == 1],
    na.rm = TRUE
  )
  n_events_nopmv_matched <- sum(
    matched_df[[death_outcome]][matched_df$prolonged_mec_vent == 0],
    na.rm = TRUE
  )

  if (n_events_pmv_matched < 5 || n_events_nopmv_matched < 5) {
    warning(sprintf(
      "Group '%s': insufficient events in matched sample (PMV=%d, no-PMV=%d). Skipping.",
      group_label, n_events_pmv_matched, n_events_nopmv_matched))
    return(NULL)
  }

  balance_obj <- summary(match_obj, standardize = TRUE)
  smd_before <- as.data.frame(balance_obj$sum.all)[, "Std. Mean Diff.", drop = FALSE]
  smd_after  <- as.data.frame(balance_obj$sum.matched)[, "Std. Mean Diff.", drop = FALSE]

  smd_table <- data.frame(
    variable   = rownames(smd_before),
    SMD_before = round(smd_before[, 1], 3),
    SMD_after  = round(smd_after[, 1], 3),
    balanced   = abs(smd_after[, 1]) < 0.1,
    stringsAsFactors = FALSE
  )
  rownames(smd_table) <- NULL

  outcome_formula <- as.formula(
    paste(death_outcome, "~ prolonged_mec_vent +", paste(covariates, collapse = " + "))
  )

  model_matched <- glm(
    formula = outcome_formula,
    data    = matched_df,
    family  = binomial(),
    weights = matched_df$weights
  )

  robust_vcov    <- vcovCL(model_matched, cluster = ~subclass, type = "HC3")
  robust_results <- coeftest(model_matched, vcov = robust_vcov)

  coef_names <- rownames(robust_results)
  estimates  <- robust_results[, "Estimate"]
  se_robust  <- robust_results[, "Std. Error"]
  pvals      <- robust_results[, "Pr(>|z|)"]

  or_psm_full <- data.frame(
    term     = coef_names,
    OR       = round(exp(estimates), 2),
    OR_lower = round(exp(estimates - 1.96 * se_robust), 2),
    OR_upper = round(exp(estimates + 1.96 * se_robust), 2),
    p.value  = round(pvals, 4),
    stringsAsFactors = FALSE
  )
  rownames(or_psm_full) <- NULL

  pmv_row_index <- which(or_psm_full$term == "prolonged_mec_vent")
  or_pmv_row <- data.frame(
    group           = group_label,
    outcome         = death_outcome,
    n_pairs         = n_matched,
    n_pmv_unmatched = n_unmatched,
    OR              = ifelse(length(pmv_row_index) == 0, NA_real_, or_psm_full$OR[pmv_row_index]),
    OR_lower        = ifelse(length(pmv_row_index) == 0, NA_real_, or_psm_full$OR_lower[pmv_row_index]),
    OR_upper        = ifelse(length(pmv_row_index) == 0, NA_real_, or_psm_full$OR_upper[pmv_row_index]),
    p.value         = ifelse(length(pmv_row_index) == 0, NA_real_, or_psm_full$p.value[pmv_row_index]),
    stringsAsFactors = FALSE
  )

  love_plot <- cobalt::love.plot(
    match_obj,
    stats      = "mean.diffs",
    thresholds = c(m = 0.1),
    abs        = TRUE,
    binary     = "std",
    title      = sprintf("Covariate Balance: %s", group_label),
    colors     = c("red3", "steelblue"),
    shapes     = c("triangle", "circle"),
    labels     = TRUE
  )

  safe_label <- gsub("[^a-z0-9]", "_", tolower(group_label))

  write.csv(
    smd_table,
    file.path(root_dir, sprintf("data/eICU/psm_balance_%s_%s.csv", safe_label, death_outcome)),
    row.names = FALSE)

  write.csv(
    or_psm_full,
    file.path(root_dir, sprintf("data/eICU/psm_or_full_%s_%s.csv", safe_label, death_outcome)),
    row.names = FALSE)

  ggsave(
    file.path(root_dir, sprintf("figures/eICU/PSM/loveplot_%s_%s.png", safe_label, death_outcome)),
    plot = love_plot, width = 7, height = 4, dpi = 300)

  return(list(
    match_obj       = match_obj,
    matched_df      = matched_df,
    smd_table       = smd_table,
    or_psm_full     = or_psm_full,
    or_pmv          = or_pmv_row,
    love_plot       = love_plot,
    n_pairs         = n_matched,
    n_pmv_unmatched = n_unmatched
  ))
}

psm_group_labels <- c("non-trauma", "non-TBI trauma", "non-severe TBI", "severe TBI")

for (sg in psm_group_labels) {
  df_sg <- final_dataset_eICU %>% filter(trauma_type_anyseq == sg)
  cat("\n---", sg, "---\n")
  cat("PMV:                    ", sum(df_sg$prolonged_mec_vent == 1, na.rm = TRUE), "\n")
  cat("no-PMV:                 ", sum(df_sg$prolonged_mec_vent == 0, na.rm = TRUE), "\n")
  cat("Hospital deaths in PMV: ", sum(df_sg$inhospital_death[df_sg$prolonged_mec_vent == 1], na.rm = TRUE), "\n")
}

df_nontrauma_eicu    <- final_dataset_eICU %>% filter(trauma_type_anyseq == "non-trauma")
df_nontbi_eicu       <- final_dataset_eICU %>% filter(trauma_type_anyseq == "non-TBI trauma")
df_nonseveretbi_eicu <- final_dataset_eICU %>% filter(trauma_type_anyseq == "non-severe TBI")
df_severetbi_eicu    <- final_dataset_eICU %>% filter(trauma_type_anyseq == "severe TBI")

psm_hospdeath_nontrauma_eicu    <- run_psm_analysis_eicu(df_nontrauma_eicu,    "inhospital_death", "non-trauma",     root_dir = root_dir)
psm_hospdeath_nontbi_eicu       <- run_psm_analysis_eicu(df_nontbi_eicu,       "inhospital_death", "non-TBI trauma", root_dir = root_dir)
psm_hospdeath_nonseveretbi_eicu <- run_psm_analysis_eicu(df_nonseveretbi_eicu, "inhospital_death", "non-severe TBI", root_dir = root_dir)
psm_hospdeath_severetbi_eicu    <- run_psm_analysis_eicu(df_severetbi_eicu,    "inhospital_death", "severe TBI",     root_dir = root_dir)
```

## Comparative table
```{r}
build_comparison_table_with_ci_eicu <- function(psm_list,
                                                original_or_list,
                                                group_labels,
                                                death_outcome,
                                                root_dir) {

  group_order <- c("non-trauma", "non-TBI trauma", "non-severe TBI", "severe TBI")

  psm_rows <- bind_rows(lapply(psm_list, function(x) {
    if (is.null(x)) return(NULL)
    x$or_pmv %>%
      mutate(
        analysis = "PSM (matched)",
        OR_CI    = sprintf("%.2f [%.2f\u2013%.2f]", OR, OR_lower, OR_upper),
        p_fmt    = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value))
      )
  }))

  original_rows <- bind_rows(mapply(function(df_or, label) {
    pval <- df_or$p.value
    data.frame(
      group           = label,
      outcome         = death_outcome,
      n_pairs         = NA_integer_,
      n_pmv_unmatched = NA_integer_,
      OR              = df_or$OR,
      OR_lower        = df_or$OR_lower,
      OR_upper        = df_or$OR_upper,
      p.value         = pval,
      analysis        = "Multivariable (original)",
      OR_CI           = sprintf("%.2f [%.2f\u2013%.2f]",
                                df_or$OR, df_or$OR_lower, df_or$OR_upper),
      p_fmt           = ifelse(pval < 0.001, "<0.001", sprintf("%.3f", pval)),
      stringsAsFactors = FALSE
    )
  }, original_or_list, group_labels, SIMPLIFY = FALSE))

  comparison <- bind_rows(original_rows, psm_rows) %>%
    mutate(group = factor(group, levels = group_order)) %>%
    arrange(group, analysis) %>%
    select(group, analysis, n_pairs, OR_CI, p_fmt) %>%
    rename(
      "Subgroup"    = group,
      "Analysis"    = analysis,
      "Pairs (PSM)" = n_pairs,
      "OR [95% CI]" = OR_CI,
      "p-value"     = p_fmt
    )

  cat(sprintf("\n========== Comparison table with CI: %s (eICU) ==========\n", death_outcome))
  print(comparison, row.names = FALSE)

  print(htmlTable(
    comparison,
    caption = sprintf(
      "Sensitivity analysis: Multivariable vs PSM \u2014 OR [95%% CI] for PMV on %s. eICU",
      death_outcome)))

  write.csv(
    comparison,
    file.path(root_dir, sprintf("data/eICU/psm_vs_multivariable_ci_%s.csv", death_outcome)),
    row.names = FALSE)

  write.csv(
    bind_rows(original_rows, psm_rows) %>% arrange(group, analysis),
    file.path(root_dir, sprintf("data/eICU/psm_vs_multivariable_%s.csv", death_outcome)),
    row.names = FALSE)

  return(invisible(comparison))
}

comparison_ci_hospdeath_eicu <- build_comparison_table_with_ci_eicu(
  psm_list = list(
    psm_hospdeath_nontrauma_eicu,
    psm_hospdeath_nontbi_eicu,
    psm_hospdeath_nonseveretbi_eicu,
    psm_hospdeath_severetbi_eicu
  ),
  original_or_list = list(
    or_hospdeath_PMV_nontrauma_eICU_multivariate,
    or_hospdeath_PMV_traumanontbi_eICU_multivariate,
    or_hospdeath_PMV_tbinonsevere_eICU_multivariate,
    or_hospdeath_PMV_tbisevere_eICU_multivariate
  ),
  group_labels  = psm_group_labels,
  death_outcome = "inhospital_death",
  root_dir      = root_dir
)
```

## Export PSM Supplementary Tables
```{r}
make_comparison_flextable_eicu <- function(df, title_label) {
  df_fmt <- df %>%
    mutate(`Pairs (PSM)` = ifelse(is.na(`Pairs (PSM)`), "\u2014",
                                  as.character(`Pairs (PSM)`)))

  flextable(df_fmt) %>%
    width(j = "Subgroup",    width = 1.5) %>%
    width(j = "Analysis",    width = 1.8) %>%
    width(j = "Pairs (PSM)", width = 0.9) %>%
    width(j = "OR [95% CI]", width = 1.5) %>%
    width(j = "p-value",     width = 0.8) %>%
    bold(part = "header") %>%
    bg(part = "header", bg = "#2c3e50") %>%
    color(part = "header", color = "white") %>%
    bg(i = seq(2, nrow(df_fmt), by = 2), bg = "#f2f2f2", part = "body") %>%
    bold(i = which(df_fmt$Analysis == "Multivariable (original)"), part = "body") %>%
    font(fontname = "Arial", part = "all") %>%
    fontsize(size = 10, part = "all") %>%
    fontsize(size = 11, part = "header") %>%
    align(j = c("Pairs (PSM)", "OR [95% CI]", "p-value"), align = "center", part = "all") %>%
    align(j = c("Subgroup", "Analysis"), align = "left", part = "all") %>%
    border_outer(part = "all", fp_border(color = "#2c3e50", width = 1.5)) %>%
    border_inner_h(part = "body", fp_border(color = "#cccccc", width = 0.5)) %>%
    set_caption(caption = title_label, style = "Table Caption") %>%
    merge_v(j = "Subgroup") %>%
    valign(j = "Subgroup", valign = "top", part = "body")
}

make_balance_flextable_eicu <- function(psm_list, group_labels, outcome_label) {
  group_order <- c("non-trauma", "non-TBI trauma", "non-severe TBI", "severe TBI")

  balance_rows <- bind_rows(mapply(function(psm_obj, label) {
    if (is.null(psm_obj)) {
      return(data.frame(
        subgroup   = label,
        variable   = "\u2014 not performed",
        SMD_before = NA_real_,
        SMD_after  = NA_real_,
        balanced   = NA,
        stringsAsFactors = FALSE
      ))
    }

    psm_obj$smd_table %>%
      mutate(subgroup = label) %>%
      filter(variable != "distance") %>%
      select(subgroup, variable, SMD_before, SMD_after, balanced)
  }, psm_list, group_labels, SIMPLIFY = FALSE))

  df_fmt <- balance_rows %>%
    mutate(
      subgroup = factor(subgroup, levels = group_order),
      variable = case_when(
        variable == "age"        ~ "Age",
        variable == "sexMale"    ~ "Sex (male)",
        variable == "sexFemale"  ~ "Sex (female)",
        variable == "sexUnknown" ~ "Sex (unknown)",
        variable == "aps"        ~ "APS",
        variable == "charlson"   ~ "Charlson",
        TRUE                     ~ variable
      ),
      balanced = ifelse(is.na(balanced), "\u2014", ifelse(balanced, "Yes", "No"))
    ) %>%
    arrange(subgroup) %>%
    select(subgroup, variable, SMD_before, SMD_after, balanced) %>%
    rename(
      "Subgroup"             = subgroup,
      "Variable"             = variable,
      "SMD before"           = SMD_before,
      "SMD after"            = SMD_after,
      "Balanced (|SMD|<0.1)" = balanced
    )

  flextable(df_fmt) %>%
    width(j = "Subgroup",             width = 1.4) %>%
    width(j = "Variable",             width = 1.4) %>%
    width(j = "SMD before",           width = 1.0) %>%
    width(j = "SMD after",            width = 1.0) %>%
    width(j = "Balanced (|SMD|<0.1)", width = 1.2) %>%
    bold(part = "header") %>%
    bg(part = "header", bg = "#2c3e50") %>%
    color(part = "header", color = "white") %>%
    bg(i = seq(2, nrow(df_fmt), by = 2), bg = "#f2f2f2", part = "body") %>%
    bg(i = which(df_fmt$`Balanced (|SMD|<0.1)` == "No"), bg = "#fde8e8", part = "body") %>%
    font(fontname = "Arial", part = "all") %>%
    fontsize(size = 10, part = "all") %>%
    fontsize(size = 11, part = "header") %>%
    align(j = c("SMD before", "SMD after", "Balanced (|SMD|<0.1)"),
          align = "center", part = "all") %>%
    border_outer(part = "all", fp_border(color = "#2c3e50", width = 1.5)) %>%
    border_inner_h(part = "body", fp_border(color = "#cccccc", width = 0.5)) %>%
    set_caption(
      caption = sprintf("Covariate balance (SMD) before and after PSM \u2014 %s. eICU",
                        outcome_label),
      style = "Table Caption") %>%
    merge_v(j = "Subgroup") %>%
    valign(j = "Subgroup", valign = "top", part = "body")
}

ft_hospdeath_eicu <- make_comparison_flextable_eicu(
  as.data.frame(comparison_ci_hospdeath_eicu),
  "Supplementary Table. Sensitivity analysis: Multivariable vs PSM \u2014 OR for PMV on hospital mortality. eICU."
)

ft_balance_hospdeath_eicu <- make_balance_flextable_eicu(
  psm_list = list(
    psm_hospdeath_nontrauma_eicu,
    psm_hospdeath_nontbi_eicu,
    psm_hospdeath_nonseveretbi_eicu,
    psm_hospdeath_severetbi_eicu
  ),
  group_labels  = psm_group_labels,
  outcome_label = "hospital mortality"
)

psm_docx_dir_eicu <- file.path(root_dir, "data/eICU/tables")
dir.create(psm_docx_dir_eicu, recursive = TRUE, showWarnings = FALSE)

doc_eicu <- read_docx() %>%
  body_set_default_section(
    prop_section(
      page_size    = page_size(width = 8.3, height = 11.7, orient = "portrait"),
      page_margins = page_mar(top = 1, bottom = 1, left = 1, right = 1)
    )
  ) %>%
  body_add_par(
    "Supplementary Material \u2014 PSM Sensitivity Analysis (eICU)",
    style = "heading 1"
  ) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_hospdeath_eicu) %>%
  body_add_break() %>%
  body_add_flextable(ft_balance_hospdeath_eicu)

output_path_eicu <- file.path(psm_docx_dir_eicu, "supplementary_PSM_tables_eICU.docx")
output_path_eicu <- tryCatch(
  {
    print(doc_eicu, target = output_path_eicu)
    output_path_eicu
  },
  error = function(e) {
    if (!grepl(" is open", conditionMessage(e), fixed = TRUE)) {
      stop(e)
    }

    fallback_path <- sub(
      "\\.docx$",
      paste0("_", format(Sys.time(), "%Y%m%d_%H%M%S"), ".docx"),
      output_path_eicu
    )

    warning(sprintf(
      "Target docx is open and could not be overwritten. Saving a timestamped copy instead: %s",
      fallback_path
    ))
    print(doc_eicu, target = fallback_path)
    fallback_path
  }
)
cat(sprintf("Saved: %s\n", output_path_eicu))
```

## Comparative Forest Plot
```{r}
original_or_hospdeath_eicu <- bind_rows(
  or_hospdeath_PMV_nontrauma_eICU_multivariate    %>% mutate(group = "non-trauma"),
  or_hospdeath_PMV_traumanontbi_eICU_multivariate %>% mutate(group = "non-TBI trauma"),
  or_hospdeath_PMV_tbinonsevere_eICU_multivariate %>% mutate(group = "non-severe TBI"),
  or_hospdeath_PMV_tbisevere_eICU_multivariate    %>% mutate(group = "severe TBI")
)

build_comparison_forest_plot_eicu <- function(original_or_df,
                                              psm_list,
                                              group_labels,
                                              death_outcome,
                                              plot_title,
                                              x_limits = c(0, 6),
                                              root_dir) {

  psm_rows <- bind_rows(mapply(function(psm_obj, label) {
    if (is.null(psm_obj)) {
      data.frame(group = label, OR = NA_real_, OR_lower = NA_real_, OR_upper = NA_real_,
                 analysis = "PSM (matched)", stringsAsFactors = FALSE)
    } else {
      psm_obj$or_pmv %>% mutate(analysis = "PSM (matched)") %>%
        select(group, OR, OR_lower, OR_upper, analysis)
    }
  }, psm_list, group_labels, SIMPLIFY = FALSE))

  plot_df <- bind_rows(
    original_or_df %>% mutate(analysis = "Multivariable (original)") %>%
      select(group, OR, OR_lower, OR_upper, analysis),
    psm_rows
  ) %>%
    mutate(
      group = factor(group, levels = rev(group_labels)),
      analysis = factor(analysis, levels = c("Multivariable (original)", "PSM (matched)")),
      direction = case_when(
        OR_lower > 1 & OR_upper > 1 ~ "Positive Association",
        OR_lower < 1 & OR_upper < 1 ~ "Negative Association",
        TRUE ~ "Inconclusive"
      )
    )

  fp <- ggplot(plot_df, aes(x = OR, xmin = OR_lower, xmax = OR_upper,
                            y = group, color = direction, shape = analysis)) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray40") +
    geom_errorbarh(height = 0.15, position = position_dodge(width = 0.5), na.rm = TRUE) +
    geom_point(size = 2.5, position = position_dodge(width = 0.5), na.rm = TRUE) +
    scale_color_manual(
      values = c("Positive Association" = "#c0392b",
                 "Negative Association" = "#27ae60",
                 "Inconclusive" = "gray70"),
      name = "Effect on Outcome",
      drop = FALSE
    ) +
    scale_shape_manual(
      values = c("Multivariable (original)" = 16, "PSM (matched)" = 17),
      name = "Analysis",
      na.translate = FALSE
    ) +
    scale_x_continuous(limits = x_limits) +
    labs(x = "Odds Ratio [95% CI]", y = NULL, title = plot_title) +
    theme_minimal() +
    theme(
      plot.background = element_rect(fill = "white", color = NA),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      plot.title = element_text(hjust = 0.5, size = 11),
      legend.position = "top",
      legend.box = "vertical",
      axis.ticks = element_blank(),
      axis.text.x = element_text(color = "#2c3e50"),
      axis.text.y = element_text(color = "#2c3e50")
    )

  safe_outcome <- gsub("_", "", death_outcome)
  ggsave(
    file.path(root_dir, sprintf("figures/eICU/PSM/fp_comparison_psm_%s.png", safe_outcome)),
    plot = fp, width = 10, height = 4, dpi = 300
  )

  return(fp)
}

fp_comparison_hospdeath_eicu <- build_comparison_forest_plot_eicu(
  original_or_df = original_or_hospdeath_eicu,
  psm_list = list(
    psm_hospdeath_nontrauma_eicu,
    psm_hospdeath_nontbi_eicu,
    psm_hospdeath_nonseveretbi_eicu,
    psm_hospdeath_severetbi_eicu
  ),
  group_labels  = psm_group_labels,
  death_outcome = "inhospital_death",
  plot_title    = "d. eICU. Outcome: Hospital Mortality",
  x_limits      = c(0, 6),
  root_dir      = root_dir
)

fp_comparison_hospdeath_eicu
```



