---
title: "03_modeling"
# output: html_document
---

# Environment

```{r}
# Libraries
library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(rlang)
library(randomForest)
library(caret)
library(kableExtra)
library(magick)
library(gridExtra)
library(webshot2)

# Function and variable creation
## Define the root directory
root_dir <- "C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent"

## Function for recategorizing column to a specific variable 
recategorize_function <- function(df, df_column, column_value) {
  df_mod <- df %>%
    mutate(!!df_column := as.factor(.data[[df_column]]),
           !!df_column := relevel(as.factor(.data[[df_column]]), ref = column_value))
  return(df_mod)
}

## Function for extracting OR
extract_coef_pval_OR <- function(model) {
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_upper = round(exp(estimate + 1.96 * std.error),2)))
} 

## Function for transforming names from OR's table and add adverse_protective variable
or_transformation <- function(df, custom_terms, variable_index) {
  df_transformed <- df %>%
    mutate(
      term = case_when(
        term == "(Intercept)" ~ "Intercept",
        TRUE ~ coalesce(custom_terms[term], term)  # Custom terms mapping or keep original
      ),
      adverse_protective = case_when(
        OR_lower > 1 & OR_upper > 1 ~ "Adverse",
        OR_lower < 1 & OR_upper < 1 ~ "Protective",
        TRUE ~ "Inconclusive"
      ),
      adverse_protective = factor(
        adverse_protective,
        levels = c("Adverse", "Protective", "Inconclusive")
      )
    ) %>%
    arrange(match(term, variable_index)) %>%
    filter(term != "Intercept")
  
  rownames(df) <- NULL #resets index
  
  return(df_transformed)
}


## Dictionary to relabel terms in the final tables (for better readability and presentation) 
dict_terms <- c(
  "prolonged_mec_vent" = "PMV",
  "trauma_type_anyseqnon-severe TBI" = "TBI non-severe",
  "trauma_type_anyseqsevere TBI" = "TBI severe",
  "trauma_type_anyseqnon-TBI trauma" = "non-TBI trauma",
  "trauma_type_anyseqnon-trauma" = "non-trauma",
  "prolonged_mec_vent:trauma_type_anyseqsevere TBI" = "PMV + TBI non-severe",
  "prolonged_mec_vent:trauma_type_anyseqnon-severe TBI" = "PMV + TBI severe",
  "prolonged_mec_vent:trauma_type_anyseqnon-TBI trauma" = "PMV + non-TBI trauma",
  "prolonged_mec_vent:trauma_type_anyseqnon-trauma" = "PMV + non-trauma",
  "anytrauma_anyseq" = "Any Trauma", 
  "prolonged_mec_vent:anytrauma_anyseq" = "PMV + Any Trauma",
  "age" = "Age (years)",
  "sexM" = "Sex (male)",
  "apsiii" = "APSIII",
  "charlson" = "Charlson", 
  "PMV for non-Trauma" = "PMV - non-trauma",
  "PMV for Trauma not TBI" = "PMV - non-TBI trauma",
  "PMV for TBI non-severe" = "PMV - non-severe TBI",
  "PMV for TBI severe" = "PMV - severe TBI"
)

## List of comorbidities
comorbidities <- c(
  "myocardial_infarct", "congestive_heart_failure", "peripheral_vascular_disease", 
  "cerebrovascular_disease", "dementia", "chronic_pulmonary_disease", 
  "rheumatic_disease", "peptic_ulcer_disease", "mild_liver_disease", 
  "diabetes_without_cc", "diabetes_with_cc", "paraplegia", "renal_disease", 
  "malignant_cancer", "severe_liver_disease", "metastatic_solid_tumor", "aids"
)

## Variable order in OR table
variable_order <- c(
  "Intercept", "Age (years)", "Sex (male)", "APSIII", "Charlson", "PMV", "non-trauma", "non-TBI trauma", "TBI non-severe", "TBI severe", "PMV + non-trauma", "PMV + non-TBI trauma", "PMV + TBI non-severe", "PMV + TBI severe", "PMV - severe TBI", "PMV - non-severe TBI", "PMV - non-TBI trauma", "PMV - non-trauma")

## Variables for table output in console
var_html <- c("term", "p.value", "OR","OR_lower","OR_upper")

## Function for performing univariate analysis
univariate_analysis <- function(df, death_outcome, df_column=NULL, output_name) {
  ## If df_column is not specified, it takes the whole dataset
  if (is.null(df_column)) {
    model <- glm(df[[death_outcome]] ~ prolonged_mec_vent,
                 data = df,
                 family = binomial())
  } else {
    model <- glm(df[[death_outcome]] ~ prolonged_mec_vent * df[[df_column]],
                 data = df,
                 family = binomial())
  }
  or_univariate <- extract_coef_pval_OR(model)
  row_name <- or_univariate[or_univariate[, 1] == "prolonged_mec_vent", var_html]
  row_name[, 1] <- output_name
  html_table_output <- htmlTable(row_name[,var_html], caption = paste("Univariate Analysis. Odds Ratio and Confidence Intervals for factors associated with", death_outcome, "- MIMIC-IV Database"))
  print(html_table_output)
  return(row_name)
}

## Function to obtain counts and mortality percentages for different columns and categories
create_summary_table <- function(df, death_outcome, df_column=NULL, row_label, column_value=NULL) {
  
  ## If df_column is not specified, it takes the whole dataset. If it is, a subset is taken
  if (is.null(df_column)) {
    subset_trauma <- df
  } else {
    subset_trauma <- subset(df, df[[df_column]] == column_value)
  }
  
  ## Filters dataframe for specific column and trauma type  
  subset_trauma_pmv <- subset(subset_trauma, prolonged_mec_vent == 1)
  subset_trauma_nonpmv <- subset(subset_trauma, prolonged_mec_vent == 0)
  
  ## Calculate totals and percentages
  total_number <- nrow(subset_trauma)
  Pct_category <- (total_number / nrow(df)) * 100
  n_PMV <- nrow(subset_trauma_pmv)
  n_not_pmv <- nrow(subset_trauma_nonpmv)
  
  Pct_PMV <- (n_PMV / total_number) * 100
  Pct_not_PMV <- (n_not_pmv / total_number) * 100
  
  ## Create summary table  
  summary_table <- data.frame(
    trauma_type = row_label,
    total_number = total_number,
    Pct_category = Pct_category,
    n_PMV = n_PMV,
    Pct_PMV = Pct_PMV, 
    mortality_pmv = (sum(subset_trauma_pmv[[death_outcome]] == 1) / n_PMV) * 100,
    n_not_pmv = n_not_pmv,
    Pct_not_PMV = Pct_not_PMV,
    mortality_not_pmv = (sum(subset_trauma_nonpmv[[death_outcome]] == 1) / n_not_pmv) * 100
  )
  
  return(summary_table)
}

# Titles for html outputs
caption_hospdeath_multivariable <- "Multivariable Analysis. Odds Ratio and Confidence Intervals for Hospital Mortality-associated Factors. MIMIC-IV Database"
caption_deathyear_multivariable <- "Multivariable Analysis. Odds Ratio and Confidence Intervals for 1 Year-Mortality-associated Factors. MIMIC-IV Database"

```

# Load File

```{r}
final_dataset_MIMICIV <- readRDS("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/mimic-iv/final_dataset_MIMICIV.rds")
```

# Hospital Death

## Total Population

```{r}
# Multivariate analysis
## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_MIMICIV,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_allpopulation_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_allpopulation_mimiciv_multivariate[, 1] <- "PMV for All Population"

## Save OR table and print in console
write.csv(or_2, file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-allpopulation_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_allpopulation_mimiciv_univariate <- univariate_analysis(df=final_dataset_MIMICIV, death_outcome="inhospital_death", output_name="PMV for All Population")
```

## non_trauma

```{r}
# Multivariate analysis
# Recategorize "non-trauma" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-trauma")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +",
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_nontrauma_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_nontrauma_mimiciv_multivariate[, 1] <- "PMV for non-Trauma"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-nontrauma_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_nontrauma_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for non-Trauma")
```

## Traumatized

```{r}
# Multivariate analysis
## Recategorize 1 as reference level in anytrauma_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "anytrauma_anyseq", "1")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_anytrauma_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_anytrauma_mimiciv_multivariate[, 1] <- "PMV for Traumatized Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-anytrauma_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_anytrauma_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="anytrauma_anyseq", output_name="PMV for Traumatized Patients")

```

## trauma_non_tbi

```{r}
# Multivariate Analysis
## Recategorize "non-TBI trauma" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-TBI trauma")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_traumanontbi_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_traumanontbi_mimiciv_multivariate[, 1] <- "PMV for Trauma not TBI"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-TraumanotTBI_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_traumanottbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for Trauma not TBI")
```

## TBI

```{r}
# Multivariate analysis
## Recategorize 1 as reference level in tbi_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "tbi_anyseq", "1")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * tbi_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbi_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbi_mimiciv_multivariate[, 1] <- "PMV for TBI Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-tbi_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_tbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="tbi_anyseq", output_name="PMV for TBI Patients")
```

## tbi_non_severe

```{r}
# Multivariate Analysis
## Recategorize "non-severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-severe TBI")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)
 
## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbinonsevere_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbinonsevere_mimiciv_multivariate[, 1] <- "PMV for TBI non-severe"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-TBInonsevere_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_nonseveretbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for TBI non-severe")

```

## tbi_severe

```{r}
# Multivariate Analysis 
## Recategorize "severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "severe TBI")

## Model
print('Logistic Regression of hospital mortality')
logreg <- glm(
  as.formula(paste("inhospital_death ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)
 
## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_hospdeath_PMV_tbisevere_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_hospdeath_PMV_tbisevere_mimiciv_multivariate[, 1] <- "PMV for TBI severe"

# Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV-TBIsevere_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_hospdeath_multivariable)

# Univariate analysis
or_hospdeath_PMV_severetbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="inhospital_death", df_column="trauma_type_anyseq", output_name="PMV for TBI severe")
```

## Summary

### Multivariate Analysis

```{r}
# Obtain combined table
MIMICIV_or_hospdeath_PMV_multivariate <- rbind(
  or_hospdeath_PMV_allpopulation_mimiciv_multivariate,
  or_hospdeath_PMV_nontrauma_mimiciv_multivariate,
  or_hospdeath_PMV_anytrauma_mimiciv_multivariate, 
  or_hospdeath_PMV_traumanontbi_mimiciv_multivariate,
  or_hospdeath_PMV_tbi_mimiciv_multivariate,
  or_hospdeath_PMV_tbinonsevere_mimiciv_multivariate,
  or_hospdeath_PMV_tbisevere_mimiciv_multivariate
) %>% 
  mutate_if(is.numeric, round, 3)
rownames(MIMICIV_or_hospdeath_PMV_multivariate) <- NULL
# Save OR table and print in console
write.csv(MIMICIV_or_hospdeath_PMV_multivariate,file.path(root_dir,'data/mimic-iv/or_hospdeath_PMV_all-trauma_mimiciv.csv'))
htmlTable(MIMICIV_or_hospdeath_PMV_multivariate[,var_html], caption = "Multivariate Analysis. Odds Ratio and Confidence Intervals for factors associated with Hospital Mortality. All types of trauma. MIMIC-IV Database")

# Assign proper names to OR table
MIMICIV_or_hospdeath_PMV_multivariate_2 <- or_transformation(MIMICIV_or_hospdeath_PMV_multivariate, dict_terms, variable_order)
# Order and select OR table for building Forest Plot
MIMICIV_or_hospdeath_PMV_multivariate_3 <- MIMICIV_or_hospdeath_PMV_multivariate_2 %>% 
  filter(term %in% c("PMV - severe TBI", "PMV - non-severe TBI", "PMV - non-TBI trauma", "PMV - non-trauma")) %>% 
  mutate(term = factor(term, levels = variable_order)) %>%
  arrange(term)

# Create a data frame with all possible levels for adverse-protective-inconclusive variable
adverse_protective_levels <- data.frame(adverse_protective = c("Adverse", "Protective", "Inconclusive"))

# Build the Forest plot
fp_PMV_alltraumas_hospdeath_mimiciv <- ggplot(MIMICIV_or_hospdeath_PMV_multivariate_3, 
       aes(x = OR, xmin = OR_lower, xmax = OR_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "a. MIMIC-IV. Outcome: Hospital Mortality") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome", 
    limits = adverse_protective_levels$adverse_protective) + #establishes limits of color scale
  scale_x_continuous(limits = c(0, 4)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))
fp_PMV_alltraumas_hospdeath_mimiciv


# Save the Forest plot as a PNG file
ggsave(file.path(root_dir, 'figures/mimic-iv/Forest_plot/fp_hospdeath_PMV_all-trauma_mimiciv.png'), plot=fp_PMV_alltraumas_hospdeath_mimiciv, width = 10, height = 4, dpi = 300)
```

### Univariate Analysis

```{r}
# Build univariate analysis table
hospdeath_univariate_summary <- rbind(
  or_hospdeath_PMV_allpopulation_mimiciv_univariate,
  or_hospdeath_PMV_nontrauma_mimiciv_univariate,
  or_hospdeath_PMV_anytrauma_mimiciv_univariate,
  or_hospdeath_PMV_traumanottbi_mimiciv_univariate,
  or_hospdeath_PMV_tbi_mimiciv_univariate,
  or_hospdeath_PMV_nonseveretbi_mimiciv_univariate,
  or_hospdeath_PMV_severetbi_mimiciv_univariate
) %>% 
  mutate_if(is.numeric, round, 3)
rownames(hospdeath_univariate_summary) <- NULL
## Save OR table and print in console
write.csv(hospdeath_univariate_summary,file.path(root_dir,'data/mimic-iv/hospdeath_univariate_summary_mimiciv.csv'))
htmlTable(hospdeath_univariate_summary, caption = "Univariate Analysis for different subgroups. MIMIC-IV Database")
```

### Counts and Percentages

```{r}
# Obtain counts and mortality percentages for each trauma subset
all_summary <- create_summary_table(df=final_dataset_MIMICIV, death_outcome="inhospital_death", row_label="All selected population")
any_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "anytrauma_anyseq", "Any Trauma", 1)
non_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "trauma_type_anyseq", "non-trauma", "non-trauma")
tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "tbi_anyseq", "TBI", 1)
nontbi_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "trauma_type_anyseq", "non-TBI trauma", "non-TBI trauma")
nonsevere_tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "trauma_type_anyseq", "non-severe TBI", "non-severe TBI")
severe_tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="inhospital_death", "trauma_type_anyseq", "severe TBI", "severe TBI")

## Build counts and percentages table
hospdeath_counts_summary <- rbind(
  all_summary,
  non_trauma_summary,
  any_trauma_summary,
  nontbi_trauma_summary,
  tbi_summary,
  nonsevere_tbi_summary,
  severe_tbi_summary
) %>% 
  mutate_if(is.numeric, round, 2)
rownames(hospdeath_counts_summary) <- NULL
## Save OR table and print in console
write.csv(hospdeath_counts_summary,file.path(root_dir,'data/mimic-iv/hospdeath_counts_summary_mimiciv.csv'))
htmlTable(hospdeath_counts_summary, caption = "Counts and percentages of different subgroups. MIMIC-IV Database")
```

# Death at year

## Total Population

```{r}
# Multivariate analysis
## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_MIMICIV,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_allpopulation_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_allpopulation_mimiciv_multivariate[, 1] <- "PMV for All Population"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-allpopulation_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_allpopulation_mimiciv_univariate <- univariate_analysis(df=final_dataset_MIMICIV, death_outcome="death_year", output_name="PMV for All Population")
```

## non_trauma

```{r}
# Multivariate analysis
## Recategorize "non-trauma" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-trauma")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +",
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_nontrauma_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_nontrauma_mimiciv_multivariate[, 1] <- "PMV for non-Trauma"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-nontrauma_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_nontrauma_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="death_year", df_column="trauma_type_anyseq", output_name="PMV for non-Trauma")
```

## Traumatized

```{r}
# Multivariate analysis
## Recategorize 1 as reference level in anytrauma_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "anytrauma_anyseq", "1")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * anytrauma_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_anytrauma_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_anytrauma_mimiciv_multivariate[, 1] <- "PMV for Traumatized Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-anytrauma_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_anytrauma_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="death_year", df_column="anytrauma_anyseq", output_name="PMV for Traumatized Patients")

```

## trauma_non_tbi

```{r}
# Multivariate Analysis
## Recategorize "non-TBI trauma" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-TBI trauma")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)
 
## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_traumanontbi_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_traumanontbi_mimiciv_multivariate[, 1] <- "PMV for Trauma not TBI"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-TraumanotTBI_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_traumanottbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="death_year", df_column="trauma_type_anyseq", output_name="PMV for Trauma not TBI")
```

## TBI

```{r}
# Multivariate analysis
## Recategorize 1 as reference level in tbi_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "tbi_anyseq", "1")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * tbi_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)

## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_tbi_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_tbi_mimiciv_multivariate[, 1] <- "PMV for TBI Patients"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-tbi_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_tbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, df_column="tbi_anyseq", death_outcome="death_year", output_name="PMV for TBI Patients")
```

## tbi_non_severe

```{r}
# Multivariate Analysis
## Recategorize "non-severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "non-severe TBI")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)
 
## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_tbinonsevere_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_tbinonsevere_mimiciv_multivariate[, 1] <- "PMV for TBI non-severe"

## Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-TBInonsevere_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_nonseveretbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="death_year", df_column="trauma_type_anyseq", output_name="PMV for TBI non-severe")

```

## tbi_severe

```{r}
# Multivariate Analysis 
## Recategorize "severe TBI" as reference level in trauma_type_anyseq variable
final_dataset_recategorized <- recategorize_function(final_dataset_MIMICIV, "trauma_type_anyseq", "severe TBI")

## Model
print('Logistic Regression of mortality at 1 year')
logreg <- glm(
  as.formula(paste("death_year ~ prolonged_mec_vent * trauma_type_anyseq + age + sex + apsiii + charlson +", 
                   paste(comorbidities, collapse = " + "))),
  data = final_dataset_recategorized,
  family = binomial()
)
 
## Extract OR, assign proper labels and save PMV's OR for this specific model 
or_1 <- extract_coef_pval_OR(logreg)
or_2 <- or_transformation(or_1, dict_terms, variable_order)
or_deathyear_PMV_tbisevere_mimiciv_multivariate <- or_2[or_2[,1]=="PMV", c(1,5:9)]
or_deathyear_PMV_tbisevere_mimiciv_multivariate[, 1] <- "PMV for TBI severe"

# Save OR table and print in console
write.csv(or_2,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV-TBIsevere_mimiciv.csv'))
htmlTable(or_2[,var_html], caption = caption_deathyear_multivariable)

# Univariate analysis
or_deathyear_PMV_severetbi_mimiciv_univariate <- univariate_analysis(df=final_dataset_recategorized, death_outcome="death_year", df_column="trauma_type_anyseq", output_name="PMV for TBI severe")
```

## Summary

### Multivariate Analysis

```{r}
# Obtain combined table
MIMICIV_or_deathyear_PMV_multivariate <- rbind(
  or_deathyear_PMV_allpopulation_mimiciv_multivariate,
  or_deathyear_PMV_nontrauma_mimiciv_multivariate,
  or_deathyear_PMV_anytrauma_mimiciv_multivariate, 
  or_deathyear_PMV_traumanontbi_mimiciv_multivariate,
  or_deathyear_PMV_tbi_mimiciv_multivariate,
  or_deathyear_PMV_tbinonsevere_mimiciv_multivariate,
  or_deathyear_PMV_tbisevere_mimiciv_multivariate
) %>% 
  mutate_if(is.numeric, round, 3)
rownames(MIMICIV_or_deathyear_PMV_multivariate) <- NULL
# Save OR table and print in console
write.csv(MIMICIV_or_deathyear_PMV_multivariate,file.path(root_dir,'data/mimic-iv/or_deathyear_PMV_all-trauma_mimiciv.csv'))
htmlTable(MIMICIV_or_deathyear_PMV_multivariate[,var_html], caption = "Multivariate Analysis. Odds Ratio and Confidence Intervals for factors associated with 1 Year-Mortality. All types of trauma. MIMIC-IV Database")

# Assign proper names to OR table
MIMICIV_or_deathyear_PMV_multivariate_2 <- or_transformation(MIMICIV_or_deathyear_PMV_multivariate, dict_terms, variable_order)
# Order and select OR table for building Forest Plot
MIMICIV_or_deathyear_PMV_multivariate_3 <- MIMICIV_or_deathyear_PMV_multivariate_2 %>% 
  filter(term %in% c("PMV - severe TBI", "PMV - non-severe TBI", "PMV - non-TBI trauma", "PMV - non-trauma")) %>% 
  mutate(term = factor(term, levels = variable_order)) %>%
  arrange(term)

# Create a data frame with all possible levels for adverse-protective-inconclusive variable
adverse_protective_levels <- data.frame(adverse_protective = c("Adverse", "Protective", "Inconclusive"))

# Build the Forest plot
fp_PMV_alltraumas_deathyear_mimiciv <- ggplot(MIMICIV_or_deathyear_PMV_multivariate_3, 
       aes(x = OR, xmin = OR_lower, xmax = OR_upper, y = term)) +
  geom_point(size = 2, aes(color = adverse_protective)) +
  geom_errorbarh(height = 0.2, aes(color = adverse_protective)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  labs(x = "Odds Ratio", y = NULL, title = "b. MIMIC-IV. Outcome: 1-Year Mortality") +
  theme_minimal() +
  scale_color_manual(
    values = c("#c0392b", "#27ae60", "gray70"),
    labels = c("Adverse", "Protective", "Inconclusive"),
    name = "Effect on Outcome", 
    limits = adverse_protective_levels$adverse_protective) + #establishes limits of color scale
  scale_x_continuous(limits = c(0, 4)) +
  theme(
    plot.background = element_rect(fill = "white"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(hjust = 0.5),
    legend.position = "top",
    axis.ticks = element_blank(),
    axis.text.x = element_text(color = "#2c3e50"),
    axis.text.y = element_text(color = "#2c3e50"
    ))
fp_PMV_alltraumas_deathyear_mimiciv


# Save the Forest plot as a PNG file
ggsave(file.path(root_dir, 'figures/mimic-iv/Forest_plot/fp_deathyear_PMV_all-trauma_mimiciv.png'), plot=fp_PMV_alltraumas_deathyear_mimiciv, width = 10, height = 4, dpi = 300)

```
### Univariate Analysis

```{r}
# Build univariate analysis table
deathyear_univariate_summary <- rbind(
  or_deathyear_PMV_allpopulation_mimiciv_univariate,
  or_deathyear_PMV_nontrauma_mimiciv_univariate,
  or_deathyear_PMV_anytrauma_mimiciv_univariate,
  or_deathyear_PMV_traumanottbi_mimiciv_univariate,
  or_deathyear_PMV_tbi_mimiciv_univariate,
  or_deathyear_PMV_nonseveretbi_mimiciv_univariate,
  or_deathyear_PMV_severetbi_mimiciv_univariate
) %>% 
  mutate_if(is.numeric, round, 3)
rownames(deathyear_univariate_summary) <- NULL
## Save OR table and print in console
write.csv(deathyear_univariate_summary,file.path(root_dir,'data/mimic-iv/deathyear_univariate_summary_mimiciv.csv'))
htmlTable(deathyear_univariate_summary, caption = "Univariate Analysis for different subgroups. MIMIC-IV Database")
```

### Counts and Percentages

```{r}
# Obtain counts and mortality percentages for each trauma subset
all_summary <- create_summary_table(df=final_dataset_MIMICIV, death_outcome="death_year", row_label="All selected population")
any_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "anytrauma_anyseq", "Any Trauma", 1)
non_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "trauma_type_anyseq", "non-trauma", "non-trauma")
tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "tbi_anyseq", "TBI", 1)
nontbi_trauma_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "trauma_type_anyseq", "non-TBI trauma", "non-TBI trauma")
nonsevere_tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "trauma_type_anyseq", "non-severe TBI", "non-severe TBI")
severe_tbi_summary <- create_summary_table(final_dataset_MIMICIV, death_outcome="death_year", "trauma_type_anyseq", "severe TBI", "severe TBI")

## Build counts and percentages table
deathyear_counts_summary <- rbind(
  all_summary,
  non_trauma_summary,
  any_trauma_summary,
  nontbi_trauma_summary,
  tbi_summary,
  nonsevere_tbi_summary,
  severe_tbi_summary
) %>% 
  mutate_if(is.numeric, round, 2)
rownames(deathyear_counts_summary) <- NULL
## Save OR table and print in console
write.csv(deathyear_counts_summary,file.path(root_dir,'data/mimic-iv/deathyear_counts_summary_mimiciv.csv'))
htmlTable(deathyear_counts_summary, caption = "Counts and percentages of different subgroups. MIMIC-IV Database")
```




