---
title: "03_modeling"
# output: html_document
---

# Environment

```{r}
library(skimr)
library(mgcv)
library(parallel)
library(parglm)
library(pROC)
library(ggplot2)
library(tibble)
library(dplyr)
library(feather)
library(broom)
library(htmlTable)
library(randomForest)
library(caret)

# metric performance
precision <- function(matrix) {
  # true positive
  tp <- matrix[2, 2]
  # false positive
  fp <- matrix[1, 2]
  return (tp / (tp + fp))
}
recall <- function(matrix) {
  # true positive
  tp <- matrix[2, 2]
  # false positive
  fn <- matrix[2, 1]
  return (tp / (tp + fn))
}
extract_coef_pval_OR <- function(model) {
  
  as.data.frame(tibble(term = names(model$coefficients),
                       estimate = model$coefficients,
                       std.error = sqrt(diag(vcov(model))),
                       statistic = model$coefficients / sqrt(diag(vcov(model))),
                       p.value = 2 * (1 - pnorm(abs(statistic))),
                       OR = round(exp(estimate),2),
                       OR_CI_lower = round(exp(estimate - 1.96 * std.error),2),
                       OR_CI_upper = round(exp(estimate + 1.96 * std.error),2)
  )
  )
} 
```

#Load File

```{r}
final_dataset <- read.csv('C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/final_dataset.csv')
```


# Logistic Regression_mortality at 30 days

## Fitting the model

```{r}
print('Logistic Regression of mortality at 30 days')
print(Sys.time())
logreg_death30days <- glm(
   death_30days ~ age + gender + hospital_stay_days + icu_stay_days + tbi_seq1 + tbi_severe_seq1 + anytrauma_seq1 + tbi_anyseq + tbi_severe_anyseq + anytrauma_anyseq + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + prolonged_mec_vent + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

## Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_death30days_1 <- extract_coef_pval_OR(logreg_death30days)

write.csv(coef_pval_df_death30days_1,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/oddsratio_death30days.csv')
htmlTable(coef_pval_df_death30days_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 30 days")
```

## Forest Plot

```{r}
coef_pval_df_death30days_2 <- coef_pval_df_death30days_1 %>% 
  mutate(orden = case_when(term == "age" ~ 1,
                           term == "genderM" ~ 2,
                           term == "hospital_stay_days" ~ 3,
                           term == "icu_stay_days" ~ 4,
                           term == "tbi_seq1" ~ 5,
                           term == "tbi_severe_seq1" ~ 6,
                           term == "anytrauma_seq1" ~ 7,
                           term == "tbi_anyseq" ~ 8,
                           term == "tbi_severe_anyseq" ~ 9,
                           term == "anytrauma_anyseq" ~ 10,
                           term == "consecutive_vm_days" ~ 11,
                           term == "mv_consecutive_7days" ~ 12,
                           term == "mv_consecutive_14days" ~ 13,
                           term == "mv_consecutive_21days" ~ 14,
                           term == "total_vm_days" ~ 15,
                           term == "mv_total_7days" ~ 16,
                           term == "mv_total_14days" ~ 17,
                           term == "mv_total_21days" ~ 18,
                           term == "mv_96traq" ~ 19,
                           term == "prolonged_mec_vent" ~ 20,
                           term == "apsiii" ~ 21,
                           term == "apsiii_prob" ~ 22,
                           term == "sapsii" ~ 23,
                           term == "sapsii_prob" ~ 24,
                           term == "gcs_min" ~ 25,
                           term == "(Intercept)" ~ 26,
                           TRUE ~ 999),
         term = case_when(term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "tbi_seq1" ~ "TBI as first diagnostics",
                           term == "tbi_severe_seq1" ~ "Severe TBI as first diagnostics",
                           term == "anytrauma_seq1" ~ "Traumatism as first diagnostics",
                           term == "tbi_anyseq" ~ "TBI as any diagnostic order",
                           term == "tbi_severe_anyseq" ~ "Severe TBI as any diagnostics",
                           term == "anytrauma_anyseq" ~ "Traumatism as any diagnostic order",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "prolonged_mec_vent" ~ "PMV - Prolonged Mechanical Ventilation",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_death30days_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 22)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 30 days') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/oddsratio_death30days.png", width = 6, height = 4, dpi = 300)
```
## Model Performance

```{r eval=FALSE, include=FALSE}
y_test <- final_dataset$death_30days

predict <- predict(logreg_death_30days, final_dataset, type = 'response')

# Confusion matrix
table_mat <- table(y_test, predict > 0.90)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

# AUTOC plot
predicted_death <- as.numeric(c(predict > 0.1))
rocobj <- roc(y_test, predicted_death)

auc <- round(auc(y_test, predicted_death),2)

# Create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics <- round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric <- rownames(metrics)
metrics$metric <- toupper(metrics$metric)

names(metrics) <- c('Value', 'Metric')

model_performance <- ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
model_performance
```

# Logistic Regression_mortality at 1 year

## Fitting the model

```{r}
print('Logistic Regression of mortality at 1 year')
print(Sys.time())
logreg_deathyear <- glm(
   death_year ~ age + gender + hospital_stay_days + icu_stay_days + tbi_seq1 + tbi_severe_seq1 + anytrauma_seq1 + tbi_anyseq + tbi_severe_anyseq + anytrauma_anyseq + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + prolonged_mec_vent + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min,
  data =  final_dataset,
  family=binomial())
print(Sys.time())
```

## Odds ratio

```{r fig.height=16, fig.width=10}
coef_pval_df_deathyear_1 <- extract_coef_pval_OR(logreg_deathyear)

write.csv(coef_pval_df_deathyear_1,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/oddsratio_deathyear.csv')
htmlTable(coef_pval_df_deathyear_1[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 1 Year")
```

## Forest Plot

```{r}
coef_pval_df_deathyear_2 <- coef_pval_df_deathyear_1 %>% 
  mutate(orden = case_when(term == "age" ~ 1,
                           term == "genderM" ~ 2,
                           term == "hospital_stay_days" ~ 3,
                           term == "icu_stay_days" ~ 4,
                           term == "tbi_seq1" ~ 5,
                           term == "tbi_severe_seq1" ~ 6,
                           term == "anytrauma_seq1" ~ 7,
                           term == "tbi_anyseq" ~ 8,
                           term == "tbi_severe_anyseq" ~ 9,
                           term == "anytrauma_anyseq" ~ 10,
                           term == "consecutive_vm_days" ~ 11,
                           term == "mv_consecutive_7days" ~ 12,
                           term == "mv_consecutive_14days" ~ 13,
                           term == "mv_consecutive_21days" ~ 14,
                           term == "total_vm_days" ~ 15,
                           term == "mv_total_7days" ~ 16,
                           term == "mv_total_14days" ~ 17,
                           term == "mv_total_21days" ~ 18,
                           term == "mv_96traq" ~ 19,
                           term == "prolonged_mec_vent" ~ 20,
                           term == "apsiii" ~ 21,
                           term == "apsiii_prob" ~ 22,
                           term == "sapsii" ~ 23,
                           term == "sapsii_prob" ~ 24,
                           term == "gcs_min" ~ 25,
                           term == "(Intercept)" ~ 26,
                           TRUE ~ 999),
         term = case_when(term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "tbi_seq1" ~ "TBI as first diagnostics",
                           term == "tbi_severe_seq1" ~ "Severe TBI as first diagnostics",
                           term == "anytrauma_seq1" ~ "Traumatism as first diagnostics",
                           term == "tbi_anyseq" ~ "TBI as any diagnostic order",
                           term == "tbi_severe_anyseq" ~ "Severe TBI as any diagnostics",
                           term == "anytrauma_anyseq" ~ "Traumatism as any diagnostic order",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "prolonged_mec_vent" ~ "PMV - Prolonged Mechanical Ventilation",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_deathyear_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.1, 21)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 1 year') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))


# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/oddsratio_deathyear.png", width = 6, height = 4, dpi = 300)
```

## Model Performance

```{r eval=FALSE, include=FALSE}
y_test <- final_dataset$death_year

predict <- predict(logreg_deathyear, final_dataset, type = 'response')

# Confusion matrix
table_mat <- table(y_test, predict > 0.90)
table_mat

accuracy <- sum(diag(table_mat)) / sum(table_mat)
accuracy

prec <- precision(table_mat)
prec

rec <- recall(table_mat)
rec

f1 <- 2 * ((prec * rec) / (prec + rec))
f1

# AUTOC plot
predicted_death <- as.numeric(c(predict > 0.1))
rocobj <- roc(y_test, predicted_death)

auc <- round(auc(y_test, predicted_death),2)

# Create ROC plot
ggroc(rocobj, colour = 'steelblue', size = 2) +
  ggtitle(paste0('ROC Curve ', '(AUC = ', auc, ')'))

metrics <- round(as.data.frame(rbind(accuracy,auc)),2)
metrics$metric <- rownames(metrics)
metrics$metric <- toupper(metrics$metric)

names(metrics) <- c('Value', 'Metric')

model_performance <- ggplot(data=metrics, aes(x=Metric, y=Value)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Value), vjust=1.6, color="white", size=3.5)+
  theme_minimal()
model_performance
```

# Logistic regression model with interaction term. Mortality at 30 days 

## Interaction term PMV and TBI
```{r}
# Fit a logistic regression model with interaction term
logreg_interactive_death30days <- glm(death_30days ~ prolonged_mec_vent * tbi_anyseq + age + gender + hospital_stay_days + icu_stay_days + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min, data = final_dataset, family = binomial)

# Print the model summary
print('Model summary of logistic regression model with interaction term between PMV and TBI for death at 30 days')
summary(logreg_interactive_death30days)

# Perform an analysis of variance (ANOVA) to test the significance of the interaction
anova(logreg_interactive_death30days, test = "Chisq")

print('If the interaction is significant, we can proceed with post-hoc analyses to explore the nature of the interaction')

# Plot the interaction effect
ggplot(final_dataset, aes(x = prolonged_mec_vent, y = death_30days, color = factor(tbi_anyseq))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Prolonged Mechanical Ventilation", y = "Mortality at 30 Days", color = "TBI", title = "Interaction Effect of TBI and PMV on Death at 30 days") +
  theme_minimal()

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/interaction_pmv_tbi_30days.png", width = 6, height = 4, dpi = 300)

# Print Odds Ratio
coef_pval_df_interactive_tbianyseq_death30days <- extract_coef_pval_OR(logreg_interactive_death30days)
htmlTable(coef_pval_df_interactive_tbianyseq_death30days[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 30 days by interaction terms PMV and TBI")

# Forest Plot
coef_pval_df_interactive_tbianyseq_death30days_2 <- coef_pval_df_interactive_tbianyseq_death30days %>% 
  mutate(orden = case_when(term == "prolonged_mec_vent" ~ 1,
                           term == "tbi_anyseq" ~ 2,
                           term == "prolonged_mec_vent:tbi_anyseq" ~ 3,
                           term == "age" ~ 4,
                           term == "genderM" ~ 5,
                           term == "hospital_stay_days" ~ 6,
                           term == "icu_stay_days" ~ 7,
                           term == "consecutive_vm_days" ~ 18,
                           term == "mv_consecutive_7days" ~ 19,
                           term == "mv_consecutive_14days" ~ 20,
                           term == "mv_consecutive_21days" ~ 21,
                           term == "total_vm_days" ~ 22,
                           term == "mv_total_7days" ~ 23,
                           term == "mv_total_14days" ~ 24,
                           term == "mv_total_21days" ~ 25,
                           term == "mv_96traq" ~ 26,
                           term == "apsiii" ~ 27,
                           term == "apsiii_prob" ~ 28,
                           term == "sapsii" ~ 29,
                           term == "sapsii_prob" ~ 30,
                           term == "gcs_min" ~ 31,
                           term == "(Intercept)" ~ 32,
                           TRUE ~ 999),
         term = case_when(term == "prolonged_mec_vent" ~ "PMV",
                           term == "tbi_anyseq" ~ "TBI",
                           term == "prolonged_mec_vent:tbi_anyseq" ~ "TBI + PMV",
                           term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_interactive_tbianyseq_death30days_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 4)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 30 days. Interaction between PMV and TBI') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/forestplot_pmv_tbi_30days.png", width = 6, height = 4, dpi = 300)


```

## Interaction term PMV and trauma_type_anyseq
```{r}
# Fit a logistic regression model with interaction term
logreg_interactive_death30days_2 <- glm(death_30days ~ prolonged_mec_vent * trauma_type_anyseq + age + gender + hospital_stay_days + icu_stay_days + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min, data = final_dataset, family = binomial)

# Print the model summary
print('Model summary of logistic regression model with interaction term between PMV and trauma type for death at 30 days')
summary(logreg_interactive_death30days_2)

# Perform an analysis of variance (ANOVA) to test the significance of the interaction
anova(logreg_interactive_death30days_2, test = "Chisq")

print('If the interaction is significant, we can proceed with post-hoc analyses to explore the nature of the interaction')

# Plot the interaction effect
ggplot(final_dataset, aes(x = prolonged_mec_vent, y = death_30days, color = factor(trauma_type_anyseq))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Prolonged Mechanical Ventilation", y = "Mortality at 30 Days", color = "Type of Trauma", title = "Interaction Effect of PMV and Type of Trauma on Death at 30 days") +
  theme_minimal()

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/interaction_pmv_traumatype_30days.png", width = 6, height = 4, dpi = 300)

# Print Odds Ratio
coef_pval_df_interactive_traumatype_death30days <- extract_coef_pval_OR(logreg_interactive_death30days_2)
htmlTable(coef_pval_df_interactive_traumatype_death30days[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 30 days by interaction terms PMV and Trauma Types")

# Forest Plot
coef_pval_df_interactive_traumatype_death30days_2 <- coef_pval_df_interactive_traumatype_death30days %>% 
  mutate(orden = case_when(term == "prolonged_mec_vent" ~ 1,
                           term == "trauma_type_anyseqtbi_non_severe" ~ 2,
                           term == "trauma_type_anyseqtbi_severe" ~ 3,
                           term == "trauma_type_anyseqtrauma_non_tbi" ~ 4,
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_non_severe" ~ 5,
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_severe" ~ 6,
                           term == "prolonged_mec_vent:trauma_type_anyseqtrauma_non_tbi" ~ 7,
                           term == "age" ~ 8,
                           term == "genderM" ~ 9,
                           term == "hospital_stay_days" ~ 10,
                           term == "icu_stay_days" ~ 11,
                           term == "consecutive_vm_days" ~ 12,
                           term == "mv_consecutive_7days" ~ 13,
                           term == "mv_consecutive_14days" ~ 14,
                           term == "mv_consecutive_21days" ~ 15,
                           term == "total_vm_days" ~ 16,
                           term == "mv_total_7days" ~ 17,
                           term == "mv_total_14days" ~ 18,
                           term == "mv_total_21days" ~ 19,
                           term == "mv_96traq" ~ 20,
                           term == "apsiii" ~ 21,
                           term == "apsiii_prob" ~ 22,
                           term == "sapsii" ~ 23,
                           term == "sapsii_prob" ~ 24,
                           term == "gcs_min" ~ 25,
                           term == "(Intercept)" ~ 26,
                           TRUE ~ 999),
         term = case_when(term == "prolonged_mec_vent" ~ "PMV",
                           term == "trauma_type_anyseqtbi_non_severe" ~ "TBI-non severe",
                           term == "trauma_type_anyseqtbi_severe" ~ "TBI-severe",
                           term == "trauma_type_anyseqtrauma_non_tbi" ~ "Trauma-not TBI",
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_non_severe" ~ "PMV + TBI-non severe",
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_severe" ~ "PMV + TBI-severe",
                           term == "prolonged_mec_vent:trauma_type_anyseqtrauma_non_tbi" ~ "PMV + Trauma-not TBI",
                           term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_interactive_traumatype_death30days_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 26)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 30 days. Interaction between PMV and Trauma Types') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/forestplot_pmv_traumatype_30days.png", width = 6, height = 4, dpi = 300)

```

## TBI and non-TBI groups
```{r}
# Conduct separate analyses for TBI and non-TBI groups
final_dataset_tbi <- subset(final_dataset, tbi_seq1 == 1)
final_dataset_notbi <- subset(final_dataset, tbi_seq1 == 0)

# Fit logistic regression models in each group
logreg_death30days_tbi <- glm(death_30days ~ age + prolonged_mec_vent + sapsii, data = final_dataset_tbi, family = binomial)
logreg_death30days_notbi <- glm(death_30days ~ age + prolonged_mec_vent + sapsii, data = final_dataset_notbi, family = binomial)

# Print the model summaries
print('Model summary for death at 30 days in the TBI group')
summary(logreg_death30days_tbi)
print('Model summary for death at 30 days in the non-TBI group')
summary(logreg_death30days_notbi)

# Odds Ratio
coef_pval_df_death30days_tbi <- extract_coef_pval_OR(logreg_death30days_tbi)
coef_pval_df_death30days_notbi <- extract_coef_pval_OR(logreg_death30days_notbi)

# Write and print table with odds ratio
write.csv(coef_pval_df_death30days_tbi,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/coef_pval_df_death30days_tbi.csv')
htmlTable(coef_pval_df_death30days_tbi[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio of dataset with TBI")

write.csv(coef_pval_df_death30days_notbi,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/coef_pval_df_death30days_notbi.csv')
htmlTable(coef_pval_df_death30days_notbi[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio of dataset without TBI")
```

# Logistic regression model with interaction term. Mortality at 1 year 

## Interaction term PMV and TBI
```{r}
# Fit a logistic regression model with interaction term
logreg_interactive_deathyear <- glm(death_30days ~ prolonged_mec_vent * tbi_anyseq + age + gender + hospital_stay_days + icu_stay_days + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min, data = final_dataset, family = binomial)

# Print the model summary
print('Model summary of logistic regression model with interaction term between PMV and TBI for death at 1 year')
summary(logreg_interactive_deathyear)

# Perform an analysis of variance (ANOVA) to test the significance of the interaction
anova(logreg_interactive_deathyear, test = "Chisq")

print('If the interaction is significant, we can proceed with post-hoc analyses to explore the nature of the interaction')

# Plot the interaction effect
ggplot(final_dataset, aes(x = prolonged_mec_vent, y = death_year, color = factor(tbi_anyseq))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Prolonged Mechanical Ventilation", y = "Mortality at 1 year", color = "TBI", title = "Interaction Effect of TBI and PMV on Death at 1 year") +
  theme_minimal()

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/interaction_pmv_tbi_year.png", width = 6, height = 4, dpi = 300)

# Print Odds Ratio
coef_pval_df_interactive_tbianyseq_deathyear <- extract_coef_pval_OR(logreg_interactive_deathyear)
htmlTable(coef_pval_df_interactive_tbianyseq_deathyear[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 1 Year by interaction terms PMV and TBI")

# Forest Plot
coef_pval_df_interactive_tbianyseq_deathyear_2 <- coef_pval_df_interactive_tbianyseq_deathyear %>% 
  mutate(orden = case_when(term == "prolonged_mec_vent" ~ 1,
                           term == "tbi_anyseq" ~ 2,
                           term == "prolonged_mec_vent:tbi_anyseq" ~ 3,
                           term == "age" ~ 4,
                           term == "genderM" ~ 5,
                           term == "hospital_stay_days" ~ 6,
                           term == "icu_stay_days" ~ 7,
                           term == "consecutive_vm_days" ~ 18,
                           term == "mv_consecutive_7days" ~ 19,
                           term == "mv_consecutive_14days" ~ 20,
                           term == "mv_consecutive_21days" ~ 21,
                           term == "total_vm_days" ~ 22,
                           term == "mv_total_7days" ~ 23,
                           term == "mv_total_14days" ~ 24,
                           term == "mv_total_21days" ~ 25,
                           term == "mv_96traq" ~ 26,
                           term == "apsiii" ~ 27,
                           term == "apsiii_prob" ~ 28,
                           term == "sapsii" ~ 29,
                           term == "sapsii_prob" ~ 30,
                           term == "gcs_min" ~ 31,
                           term == "(Intercept)" ~ 32,
                           TRUE ~ 999),
         term = case_when(term == "prolonged_mec_vent" ~ "PMV",
                           term == "tbi_anyseq" ~ "TBI",
                           term == "prolonged_mec_vent:tbi_anyseq" ~ "TBI + PMV",
                           term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_interactive_tbianyseq_deathyear_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.1, 4)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 1 year. Interaction between PMV and TBI') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/forestplot_pmv_tbi_year.png", width = 6, height = 4, dpi = 300)

```

## Interaction term PMV and trauma_type
```{r}
# Fit a logistic regression model with interaction term
logreg_interactive_deathyear_2 <- glm(death_30days ~ prolonged_mec_vent * trauma_type_anyseq + age + gender + hospital_stay_days + icu_stay_days + consecutive_vm_days + mv_consecutive_7days + mv_consecutive_14days + mv_consecutive_21days + total_vm_days + mv_total_7days + mv_total_14days + mv_total_21days + mv_96traq + apsiii + apsiii_prob + sapsii + sapsii_prob + gcs_min, data = final_dataset, family = binomial)

# Print the model summary
print('Model summary of logistic regression model with interaction term between PMV and trauma type for death at 1 year')
summary(logreg_interactive_deathyear_2)



# Perform an analysis of variance (ANOVA) to test the significance of the interaction
anova(logreg_interactive_deathyear_2, test = "Chisq")

print('If the interaction is significant, we can proceed with post-hoc analyses to explore the nature of the interaction')

# Plot the interaction effect
ggplot(final_dataset, aes(x = prolonged_mec_vent, y = death_year, color = factor(trauma_type_anyseq))) +
  geom_point() +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
  labs(x = "Prolonged Mechanical Ventilation", y = "Mortality at 1 Year", color = "Type of Trauma", title = "Interaction Effect of PMV and Type of Trauma on Death at 1 Year") +
  theme_minimal()

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/interaction_pmv_traumatype_year.png", width = 6, height = 4, dpi = 300)

# Print Odds Ratio
coef_pval_df_interactive_traumatype_deathyear <- extract_coef_pval_OR(logreg_interactive_deathyear_2)
htmlTable(coef_pval_df_interactive_traumatype_deathyear[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio for death at 1 Year by interaction terms PMV and Trauma Types")

# Forest Plot
coef_pval_df_interactive_traumatype_deathyear_2 <- coef_pval_df_interactive_traumatype_deathyear %>% 
  mutate(orden = case_when(term == "prolonged_mec_vent" ~ 1,
                           term == "trauma_type_anyseqtbi_non_severe" ~ 2,
                           term == "trauma_type_anyseqtbi_severe" ~ 3,
                           term == "trauma_type_anyseqtrauma_non_tbi" ~ 4,
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_non_severe" ~ 5,
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_severe" ~ 6,
                           term == "prolonged_mec_vent:trauma_type_anyseqtrauma_non_tbi" ~ 7,
                           term == "age" ~ 8,
                           term == "genderM" ~ 9,
                           term == "hospital_stay_days" ~ 10,
                           term == "icu_stay_days" ~ 11,
                           term == "consecutive_vm_days" ~ 12,
                           term == "mv_consecutive_7days" ~ 13,
                           term == "mv_consecutive_14days" ~ 14,
                           term == "mv_consecutive_21days" ~ 15,
                           term == "total_vm_days" ~ 16,
                           term == "mv_total_7days" ~ 17,
                           term == "mv_total_14days" ~ 18,
                           term == "mv_total_21days" ~ 19,
                           term == "mv_96traq" ~ 20,
                           term == "apsiii" ~ 21,
                           term == "apsiii_prob" ~ 22,
                           term == "sapsii" ~ 23,
                           term == "sapsii_prob" ~ 24,
                           term == "gcs_min" ~ 25,
                           term == "(Intercept)" ~ 26,
                           TRUE ~ 999),
         term = case_when(term == "prolonged_mec_vent" ~ "PMV",
                           term == "trauma_type_anyseqtbi_non_severe" ~ "TBI-non severe",
                           term == "trauma_type_anyseqtbi_severe" ~ "TBI-severe",
                           term == "trauma_type_anyseqtrauma_non_tbi" ~ "Trauma-not TBI",
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_non_severe" ~ "PMV + TBI-non severe",
                           term == "prolonged_mec_vent:trauma_type_anyseqtbi_severe" ~ "PMV + TBI-severe",
                           term == "prolonged_mec_vent:trauma_type_anyseqtrauma_non_tbi" ~ "PMV + Trauma-not TBI",
                           term == "age" ~ "Age (years)",
                           term == "genderM" ~ "Gender (male)",
                           term == "hospital_stay_days" ~ "Lengh of hospital stay (days)",
                           term == "icu_stay_days" ~ "Lenght of ICU stay (days)",
                           term == "consecutive_vm_days" ~ "Consecutive days with >6 hour ventilation",
                           term == "mv_consecutive_7days" ~ "PMV > consecutive 7 days",
                           term == "mv_consecutive_14days" ~ "PMV > consecutive 14 days",
                           term == "mv_consecutive_21days" ~ "PMV > consecutive 21 days",
                           term == "total_vm_days" ~ "Total days with >6 hour ventilation",
                           term == "mv_total_7days" ~ "PMV > 7 days in total",
                           term == "mv_total_14days" ~ "PMV > 14 days in total",
                           term == "mv_total_21days" ~ "PMV > 21 days in total",
                           term == "mv_96traq" ~ "MV for 96h + traqueostomy",
                           term == "apsiii" ~ "APSIII",
                           term == "apsiii_prob" ~ "APSIII probability",
                           term == "sapsii" ~ "SAPSII",
                           term == "sapsii_prob" ~ "SAPSII probability",
                           term == "gcs_min" ~ "GCS",
                           TRUE ~ term)) %>% 
  arrange(desc(orden)) %>% 
  select(-orden) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = factor(term, levels = term, ordered = TRUE))

ggplot(coef_pval_df_interactive_traumatype_deathyear_2, 
       aes(x=term, y=OR, ymin=OR_CI_lower, ymax=OR_CI_upper)) +
  geom_hline(yintercept=1, colour = "blue", linewidth=0.8, linetype="dotted") +
  geom_pointrange(alpha=0.7, shape=15) +
  geom_vline(xintercept=0, linewidth=1.2) +
  scale_x_discrete(expand = expansion(add = c(1, 0.8))) + 
  scale_y_continuous(expand = expansion(add = c(0.1, 0.1))) +
  scale_y_log10(limits = c(0.01, 25)) +
  coord_flip() +
  labs(y="Odds Ratio", x=NULL) +
  ggtitle('Mortality at 1 Year. Interaction between PMV and Trauma Types') +
  theme_minimal() +
  theme(legend.text = element_text(size = 10), legend.spacing.y = unit(1, "cm"))

# Save the plot as a PNG file
ggsave("C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/figures/forestplot_pmv_traumatype_year.png", width = 6, height = 4, dpi = 300)

```

## TBI and non-TBI groups
```{r}
# Fit logistic regression models in each group
logreg_deathyear_tbi <- glm(death_year ~ age + prolonged_mec_vent + sapsii, data = final_dataset_tbi, family = binomial)
logreg_deathyear_notbi <- glm(death_year ~ age + prolonged_mec_vent + sapsii, data = final_dataset_notbi, family = binomial)

# Print the model summaries
print('Model summary for death at 1 year in the TBI group')
summary(logreg_deathyear_tbi)
print('Model summary for death at 1 year in the TBI group')
summary(logreg_deathyear_notbi)

# Odds Ratio
coef_pval_df_deathyear_tbi <- extract_coef_pval_OR(logreg_deathyear_tbi)
coef_pval_df_deathyear_notbi <- extract_coef_pval_OR(logreg_deathyear_notbi)

# Write and print table with odds ratio
write.csv(coef_pval_df_deathyear_tbi,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/oddsratio_deathyear_tbi.csv')
htmlTable(coef_pval_df_deathyear_tbi[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio of dataset with TBI")

write.csv(coef_pval_df_deathyear_notbi,'C:/Users/se_al/Documents/GitHub/mimiciv-mec-vent/data/oddsratio_deathyear_notbi.csv')
htmlTable(coef_pval_df_deathyear_notbi[,c("term","OR","OR_CI_lower","OR_CI_upper")], caption = "Odds Ratio of dataset without TBI")
```

# Correlation between TBI and age
```{r}
# Calculate the point-biserial correlation
correlation_tbi_age <- cor.test(final_dataset$tbi_seq1, final_dataset$age)

# Print the correlation coefficient and p-value
cat("Point-biserial correlation coefficient:", correlation_tbi_age$estimate, "\n")
cat("p-value:", correlation_tbi_age$p.value, "\n")

# Check if the correlation is significant
if (correlation_tbi_age$p.value < 0.05) {
  cat("The correlation is significant.\n")
} else {
  cat("The correlation is not significant.\n")
}

```

